---
title: "ILD Adaptation"
author: "A.P. Zweistra"
format: html
editor: visual
---

This document adapts the original code by Bailey et al. to fit models with intensive longitudinal data (ILD) with 30 waves of data. Large parts of the code will be the same, with the main changes occurring in the model specifications.

### Init

```{r}
## random seed and packages 
set.seed(427)

library(mvtnorm)
library(psych)
library(ggplot2)
library(reshape)
library(gridExtra)
library(lavaan)
library(reshape2)

# extra for me
library(tidyverse)

## Function definitions
row <- function( M ){
    if( ! ( is.matrix(M) && dim(M)[1] == dim(M)[2] && dim(M)[1]>0 ) ){
        rowM <- NULL
    } else {
        F <- dim(M)[1]
        rowM <- matrix( NA, nrow=F^2, ncol=1)
        for ( r in (1:F) ) {
            for( c in (1:F) ) {
                rowM[ c+F*(r-1), 1 ] <- M[r,c]
            }
        }
    }
    return( rowM )
}
irow <- function( rowM ){
    if( ! ( !is.null(rowM) && ( is.matrix(rowM) || (is.array(rowM) && length( which( dim(rowM) > 1 ) ) == 1) ) && dim(rowM)[1]>0 && dim(rowM)[2]==1 ) ){
        M <- NULL
    } else {
        F <- sqrt( length( rowM ) )
        M <- matrix( NA, nrow=F, ncol=F )
        for ( r in (1:F) ) {
            for ( c in (1:F) ) {
                M[r,c] <- rowM[ c+F*(r-1) ]
            }
        }
    }
    return( M )
}
```

## Additional Functions

Let's first get a function allowing us to procedurally generate RI-CLPM model specifications. Courtesy of Gemini:

```{r}
#' Generate Lavaan Syntax for a RI-CLPM
#'
#' This function creates the model specification string for a
#' Random-Intercept Cross-Lagged Panel Model (RI-CLPM) for a
#' given number of timepoints (T).
#'
#' @param T An integer specifying the total number of timepoints. Must be > 1.
#'
#' @return A single character string containing the complete lavaan model syntax.
#'
#' @examples
#' # Generate a model for 8 timepoints
#' model_t8 <- generate_riclpm(T = 8)
#' cat(model_t8)
#'
#' # Generate a model for 50 timepoints
#' model_t50 <- generate_riclpm(T = 50)
#' # View the first few lines
#' # head(strsplit(model_t50, "\n")[[1]])

generate_riclpm <- function(T) {

  # --- 1. Input Validation ---
  if (!is.numeric(T) || T <= 1 || T %% 1 != 0) {
    stop("T must be an integer greater than 1.")
  }

  # --- 2. Procedurally Generate Model Components ---

  # Random Intercepts (RIx and RIy)
  ri_x <- sprintf("RIx =~ %s", paste(sprintf("1*x%d", 1:T), collapse = " + "))
  ri_y <- sprintf("RIy =~ %s", paste(sprintf("1*y%d", 1:T), collapse = " + "))

  # Within-person component definitions (wx and wy)
  wx_defs <- paste(sprintf("wx%d =~ 1*x%d", 1:T, 1:T), collapse = "\n")
  wy_defs <- paste(sprintf("wy%d =~ 1*y%d", 1:T, 1:T), collapse = "\n")

  # Autoregressive and cross-lagged paths (from T2 to T)
  timepoints_reg <- 2:T
  lagged_timepoints <- 1:(T - 1)
  paths_wx <- paste(sprintf("wx%d ~ ax*wx%d + by*wy%d", timepoints_reg, lagged_timepoints, lagged_timepoints), collapse = "\n")
  paths_wy <- paste(sprintf("wy%d ~ bx*wx%d + ay*wy%d", timepoints_reg, lagged_timepoints, lagged_timepoints), collapse = "\n")

  # Constrained correlated residuals (from T2 to T)
  res_covs <- paste(sprintf("wx%d ~~ ur*wy%d", 2:T, 2:T), collapse = "\n")

  # Variances of within-person components (all timepoints)
  var_wx <- paste(sprintf("wx%d ~~ wx%d", 1:T, 1:T), collapse = "\n")
  var_wy <- paste(sprintf("wy%d ~~ wy%d", 1:T, 1:T), collapse = "\n")

  # Fix observed variances to zero
  zero_var_x <- paste(sprintf("x%d ~~ 0*x%d", 1:T, 1:T), collapse = "\n")
  zero_var_y <- paste(sprintf("y%d ~~ 0*y%d", 1:T, 1:T), collapse = "\n")


  # --- 3. Assemble Final Model String ---

  model_string <- paste(
    "# 1. Random Intercepts", ri_x, ri_y,
    "# 2. Within-person components", wx_defs, wy_defs,
    "# 3. Autoregressive and cross-lagged paths", paths_wx, paths_wy,
    "# 4. Covariances",
    "wx1 ~~ wy1 # Covariance at T1",
    res_covs,
    "# 5. (Co)variances of Random Intercepts",
    "RIx ~~ varRIx*RIx",
    "RIy ~~ varRIy*RIy",
    "RIx ~~ covRI*RIy",
    "# 6. (Residual) variances of within-person components",
    var_wx,
    var_wy,
    "# 7. Fix observed variances to zero",
    zero_var_x,
    zero_var_y,
    "# 8. Fix RI covariances with first state to zero",
    "wx1 ~~ 0*RIx",
    "wx1 ~~ 0*RIy",
    "wy1 ~~ 0*RIx",
    "wy1 ~~ 0*RIy",
    sep = "\n\n" # Separate sections with a blank line for readability
  )

  return(model_string)
}
```

Next up, we run the data generating mechanism up to our maximum amount of timepoints, say we take 100 as our max:

```{r}
set.seed(42)
# Set parameters for the data generation model
## Number of participants
N <- 20000 

## Number of generated time series
T <- 100

## Number of X-related confounders
c1 <- 5

## Number of Y-related confounders
c2 <- 5

## Autoregressive effects (for all variables)
alpha <- 0.2

## Cross-lagged effects of the focal effect (X and Y)
beta_focal <- 0.1

## Main confounding effects: Cross-lagged effects 
## (1) between X-related confounders and X
## (2) between Y-related confounders and Y
## (3) within X-related confounders
## (4) within Y-related confounders
## They are all set to be equal in this simulation but can be easily changed

beta_covfocal <- 0.1

## Trivial confounding effects: Cross-lagged effects 
## (1) between X-related confounders and Y
## (2) between Y-related confoudners and X
## (3) between X-related confounder and Y-related confounders
##  They are all set to be equal in this simulation but can be easily changed

beta_covnofocal <- 0.01

## Time-specific residual variance
sigma <- 0.5


# Create objects for the simulation

## total number of variables
v <- c1 + c2 + 2


## A: Matrix of path coefficients
##   A[1,2] - Effects of V2 at t-1 on V1 at t
##   A[2,1] - Effects of V1 at t-1 on V2 at t
##


A <- matrix(NA, v, v)
diag(A) <- alpha

A[lower.tri(A, diag = FALSE)] <- beta_covnofocal 
A[upper.tri(A, diag = FALSE)] <- beta_covnofocal

c1t <- 3:(2+c1)
c2t <- (3+c1):v

A[c1t, 1] <- beta_covfocal
A[1, c1t] <- beta_covfocal

A[c2t, 2] <- beta_covfocal
A[2, c2t] <- beta_covfocal


A[c1t, c1t][lower.tri(A[c1t, c1t], diag = FALSE)] <- beta_covfocal
A[c1t, c1t][upper.tri(A[c1t, c1t], diag = FALSE)] <- beta_covfocal

A[c2t, c2t][lower.tri(A[c2t, c2t], diag = FALSE)] <- beta_covfocal
A[c2t, c2t][upper.tri(A[c2t, c2t], diag = FALSE)] <- beta_covfocal

A[1,2] <- beta_focal
A[2,1] <- beta_focal




## Check the max eigenvalue (to check convergence --- see the manuscript)
# max(eigen(A)$values)

## Sigma: Var-cov matrix of time-specific residuals
Sigma <- diag(rep(sigma, v))

## Sigma1: Stationary Var-cov matrix for the first time point
Sigma1 <- irow( solve( diag(dim(A)[1]^2) - t(A) %x% t(A) ) %*% row(Sigma) )

## Mean1: Stationary means for first time point (as no process intercepts are defined/used, apparently 0)
Mean1 <- matrix( 0, nrow=dim(A)[1], ncol=1 )



# Generate data

## Wide-format "bivariate" (X and Y) data df: N x 2*T
## df stores the data from all time points
df <- matrix(NA, nrow = N, ncol = 2*T)
colnames(df) <- c(paste("x", 1:T, sep = ""), paste("y", 1:T, sep = ""))

## data at each time point D: N x v (total number of variables) 
## D stores the data for each time point only, i.e., D will be updated in a loop
## Generate the initial values for X and Y and store them in D
D <- rmvnorm(N, mean = Mean1, Sigma1)

## Store the initial data from D to df. 
## Store the initial X (the first column of D) and Y (the second column of D) in df
df[, 1] <- D[,1]
df[, 1+T] <- D[, 2]



## Update D and df till time T

for (i in 2:T){
  D <-  D %*% t(A) + rmvnorm(N, sigma = Sigma)

  df[, i] <- D[,1]
  df[, i+T] <- D[, 2]
}

```

Now we can run the actual simulation. Fit RI-CLPM models on procedurally more timepoints and extract the variance of the RI's.

```{r}
fits <- list()
varRIx <- NULL
varRIy <- NULL

for (i in seq(3,90,3)){
  modelSpec <- generate_riclpm(i)
  
  fit <- sem(model = modelSpec, data = df)
  
  fits[[i/3]] <- fit
  varRIx[i/3] <- coef(fit)["varRIx"]
  varRIy[i/3] <- coef(fit)["varRIy"]
  
  print(i)
}

#varRIx_alpha_420 <- varRIx
saveRDS(varRIx_alpha_420, file = "varRIx_alpha_420.RDS")

#fits_alpha_420 <- fits
saveRDS(fits_alpha_420, file = "fits_alpha_420.RDS")

```

```{r}
plot(y = varRIx, x = seq(3, 90, 3), 
     main = "Variance of random intercept of x by amount of timepoints",
     xlab = "Timepoints")
plot(y = varRIy, x = seq(3, 90, 3), 
     main = "Variance of random intercept of y by amount of timepoints",
     xlab = "Timepoints")
```

Plotting them together

```{r}

data.frame(
  Timepoints = seq(3, 90, 3),
  varRIx_alpha_2 = 
    readRDS("Thesis-Illusionary-Traits/data/SIM1/varRIx_alpha_2.RDS"),
  varRIx_alpha_3 = 
    readRDS("Thesis-Illusionary-Traits/data/SIM1/varRIx_alpha_3.RDS"),
  varRIx_alpha_35= 
    readRDS("Thesis-Illusionary-Traits/data/SIM1/varRIx_alpha_35.RDS"),
  varRIx_alpha_4 = 
    readRDS("Thesis-Illusionary-Traits/data/SIM1/varRIx_alpha_420.RDS"),
  varRIx_alpha_420 = 
    readRDS("Thesis-Illusionary-Traits/data/SIM1/varRIx_alpha_420.RDS")
) %>%
  pivot_longer(
    cols = 
      c("varRIx_alpha_2","varRIx_alpha_3","varRIx_alpha_35",
        "varRIx_alpha_4","varRIx_alpha_420"),
    names_to = "Autoregressive_Effects",
    values_to = "Estimated_Variance"
    ) %>%
  ggplot(aes(x = Timepoints, y = Estimated_Variance, color = Autoregressive_Effects)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Random Intercept Variance for Differing Autoregressive Effects",
    y = "Estimated Variance",
    color = "Autoregressive Effects"
  ) +
  theme_minimal()
```

Plot individual sequences for x and y

```{r}
library(reshape2)
source("functionSource.R")

testData <- DGM(timepoints = 80)
plot_xy_sequences(testData$data, 
                  timepoints = testData$parameters$timepoints, n_participants = 1)

```

Two additional smaller simulations

1.  N participants vary (300, 600, 900)
2.  N confounders vary (3, 5, 7)

Other parameters should stay the same as in Bailey i guess? Maybe check stationarity for more confounders with stronger autoregressive effects.

```{r}
source("functionSource.R")
N_part <- c(300, 600, 900)
timepoints <- 78
filename_prefix <- "N_participants_sim"
data_struct <- create_metadata()

# N participants
for (a in seq(1:3)){
  cat("Starting simulation for condition ", a, "\n")
  
  data <- DGM(N = N_part[a], timepoints = timepoints)
  
  # Initialize storage for this N condition
  condition_results <- data.frame(
    N_participants = integer(),
    timepoints_used = integer(),
    iteration = integer(),
    varRIx = numeric(),
    varRIy = numeric(),
    convergence = logical(),
    fit_measures = I(list()),  # list column
    coefficients = I(list()),  
    processing_time = numeric()
  )
  
  for (i in seq(3,timepoints,3)){
    start_time <- Sys.time()  
    
    modelSpec <- generate_riclpm(i)
    
    fit <- sem(model = modelSpec, data = data$data)
    
    converged <- lavInspect(fit, "converged")
    fit_measures <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli",
                                       "rmsea", "srmr"))
    
    coeff <- if(converged) {coef(fit)} else {NULL}  
    
    condition_results <- rbind(condition_results, data.frame(
        N_participants = N_part[a],  
        timepoints_used = i,
        iteration = i/3,
        varRIx = if(converged) coef(fit)["varRIx"] else NA,
        varRIy = if(converged) coef(fit)["varRIy"] else NA,
        convergence = converged,
        fit_measures = I(list(fit_measures)),
        coefficients = I(list(coeff)),
        processing_time = as.numeric(difftime(Sys.time(), start_time, 
                                              units = "secs"))
      ))
    
    rm(fit)
  }
  
  data_struct$results <- rbind(data_struct$results, condition_results)
  
  filename <- save_clear(data_struct, filename_prefix)
  
  cat("Completed Sim no. ", a, "\n")
  
  rm(data)
  gc()
}
```

```{r}
results <- N_participants_sim_20250902_142615$results

# Create the plot
varRix_plot <- ggplot(data = results, 
                      aes(x = timepoints_used, 
                          y = varRIx, 
                          group = factor(N_participants), 
                          color = factor(N_participants))) +
  geom_line(linewidth = 1) +  # Add the line layer
  geom_point(size = 2) +     # Add points to mark the actual data points
  
  # --- Customize labels and titles ---
  labs(title = "varRIx over Timepoints by Number of Participants",
       x = "Timepoints Used",
       y = "varRIx Value",
       color = "N Participants") + # This renames the legend title
       
  # --- Apply a clean theme ---
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16), # Center the title
    legend.position = "bottom" # Move legend to the bottom
  )

# --- Display the plot ---
print(varRix_plot)

# --- Optionally, save the plot to a file ---
ggsave("varRix_simulation_plot.png", plot = varRix_plot, width = 10, height = 7, dpi = 300)
```

Confounder variation

```{r}
source("functionSource.R")
N_conf <- c(3, 5, 7)
timepoints <- 78
filename_prefix <- "N_confounder_sim"
data_struct <- create_metadata()

# N participants
for (a in seq(1:3)){
  cat("Starting simulation for condition ", a, "\n")
  
  data <- DGM(N = 20000, n_x_confounders = N_conf[a], n_y_confounders = N_conf[a],
              timepoints = timepoints)
  
  # Initialize storage for this N condition
  condition_results <- data.frame(
    N_confounders = integer(),
    timepoints_used = integer(),
    iteration = integer(),
    varRIx = numeric(),
    varRIy = numeric(),
    convergence = logical(),
    fit_measures = I(list()),  # list column
    coefficients = I(list()),  
    processing_time = numeric()
  )
  
  for (i in seq(3,timepoints,3)){
    start_time <- Sys.time()  
    
    modelSpec <- generate_riclpm(i)
    
    fit <- sem(model = modelSpec, data = data$data)
    
    converged <- lavInspect(fit, "converged")
    fit_measures <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli",
                                       "rmsea", "srmr"))
    
    coeff <- if(converged) {coef(fit)} else {NULL}  
    
    condition_results <- rbind(condition_results, data.frame(
        N_confounders = N_conf[a],  
        timepoints_used = i,
        iteration = i/3,
        varRIx = if(converged) coef(fit)["varRIx"] else NA,
        varRIy = if(converged) coef(fit)["varRIy"] else NA,
        convergence = converged,
        fit_measures = I(list(fit_measures)),
        coefficients = I(list(coeff)),
        processing_time = as.numeric(difftime(Sys.time(), start_time, 
                                              units = "secs"))
      ))
    
    rm(fit)
  }
  
  data_struct$results <- rbind(data_struct$results, condition_results)
  
  filename <- save_clear(data_struct, filename_prefix)
  
  cat("Completed Sim no. ", a, "\n")
  
  rm(data)
  gc()
}
```

```{r}
results <- N_confounder_sim_20250902_145933$results

# Create the plot
varRix_plot <- ggplot(data = results, 
                      aes(x = timepoints_used, 
                          y = varRIx, 
                          group = factor(N_confounders), 
                          color = factor(N_confounders))) +
 # lims(y = c(0, 0.1))+ # to remove the outlying condition from the plot
  geom_line(linewidth = 1) +  # Add the line layer
  geom_point(size = 2) +     # Add points to mark the actual data points
  
  # --- Customize labels and titles ---
  labs(title = "varRIx over Timepoints by Number of Confounders",
       x = "Timepoints Used",
       y = "varRIx Value",
       color = "N Confounders") + # This renames the legend title
       
  # --- Apply a clean theme ---
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16), # Center the title
    legend.position = "bottom" # Move legend to the bottom
  )

# --- Display the plot ---
print(varRix_plot)



```

My next simulation should involve checking the interplay between the bias of the focal crosslagged effects and the random intercept variance. Bias of the crosslagged effects is defined as the estimated cross-lagged effect - the actual cross-lagged effect (0.1)

```{r}
# Extract cross-lagged coefficients and calculate bias
results <- N_confounder_sim_20250902_145933$results

# True value of cross-lagged paths from your DGM (beta_focal = 0.1)
true_beta_focal <- 0.1

# Extract bx and by coefficients and calculate bias
results$bx <- sapply(results$coefficients, function(coef_list) {
  if(!is.null(coef_list) && "bx" %in% names(coef_list)) {
    return(coef_list["bx"])
  } else {
    return(NA)
  }
})

results$by <- sapply(results$coefficients, function(coef_list) {
  if(!is.null(coef_list) && "by" %in% names(coef_list)) {
    return(coef_list["by"])
  } else {
    return(NA)
  }
})

# Calculate bias (estimated - true)
results$bias_bx <- results$bx - true_beta_focal
results$bias_by <- results$by - true_beta_focal

# Plot bias for bx (X -> Y cross-lagged path)
bias_bx_plot <- ggplot(data = results, 
                       aes(x = timepoints_used, 
                           y = bias_bx, 
                           group = factor(N_confounders), 
                           color = factor(N_confounders))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.7) +
  labs(title = "Bias in Cross-lagged Path bx (X→Y) over Timepoints by Number of Confounders",
       x = "Timepoints Used",
       y = "Bias in bx (Estimated - True)",
       color = "N Confounders") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.position = "bottom"
  )

print(bias_bx_plot)

# Plot bias for by (Y -> X cross-lagged path)
bias_by_plot <- ggplot(data = results, 
                       aes(x = timepoints_used, 
                           y = bias_by, 
                           group = factor(N_confounders), 
                           color = factor(N_confounders))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.7) +
  labs(title = "Bias in Cross-lagged Path by (Y→X) over Timepoints by Number of Confounders",
       x = "Timepoints Used",
       y = "Bias in by (Estimated - True)",
       color = "N Confounders") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    legend.position = "bottom"
  )

print(bias_by_plot)
```

Check stationarity for more confounders , max eigenvalue is still under 1 for 7 confounders on x and y. Let's also do a visual check of some sequences.

```{r}
source("functionSource.R")

df <- DGM(n_x_confounders = 7, n_y_confounders = 7, timepoints = 2000)

plot_xy_sequences(df$data, 
                  timepoints = df$parameters$timepoints, n_participants = 3)
```
