---
title: "Thesis Simulations"
author: "A.P. Zweistra"
format: html
editor: visual
---

This is intended to be the final, complete file that contains all necessary code to reproduce the results presented in the thesis (in combination with the "functionSource.R" file.

### Initialize

```{r}
library(mvtnorm)
library(tidyverse)
library(lavaan)

# Load function source file
source("functionSource.R")

## Function definitions from Bailey, 2024
row <- function( M ){
    if( ! ( is.matrix(M) && dim(M)[1] == dim(M)[2] && dim(M)[1]>0 ) ){
        rowM <- NULL
    } else {
        F <- dim(M)[1]
        rowM <- matrix( NA, nrow=F^2, ncol=1)
        for ( r in (1:F) ) {
            for( c in (1:F) ) {
                rowM[ c+F*(r-1), 1 ] <- M[r,c]
            }
        }
    }
    return( rowM )
}
irow <- function( rowM ){
    if( ! ( !is.null(rowM) && ( is.matrix(rowM) || (is.array(rowM) && length( which( dim(rowM) > 1 ) ) == 1) ) && dim(rowM)[1]>0 && dim(rowM)[2]==1 ) ){
        M <- NULL
    } else {
        F <- sqrt( length( rowM ) )
        M <- matrix( NA, nrow=F, ncol=F )
        for ( r in (1:F) ) {
            for ( c in (1:F) ) {
                M[r,c] <- rowM[ c+F*(r-1) ]
            }
        }
    }
    return( M )
}
```

### Simulation 1: the performance of the RI-CLPM over time.

2x2 matrix with DGM params nconf(4,6) and ax(.2,.3). Illustrate role of "temporal stability" in Illusory Traits. Illustrate direct relationship between illusory random intercept variance and less biased estimation of cross-lagged params. Population matrices.

```{r}
# conditions
autoreg <- c(.2, .3)
nconf <- c(4, 6)

# general settings
timepoints <- 75
seed <- 42
filename_prefix <- "Simulation1"

# create empty data structure
data_struct <- create_metadata()

# loop through conditions
for (i in autoreg) {
  for (j in nconf) {

    # Initialise per condition results
    condition_results <- data.frame(
      autoreg         = numeric(),
      nconf           = integer(),
      timepoints_used = integer(),
      iteration       = integer(),
      varRIx          = numeric(),
      varRIy          = numeric(),
      convergence     = logical(),
      fit_measures    = I(list()),
      coefficients    = I(list()),
      processing_time = numeric()
    )

    # generate data
    data <- DGM(autocorr_effects  = i,
                n_x_confounders   = j,
                n_y_confounders   = j,
                timepoints        = timepoints,
                seed              = seed)

    # fit models on multiples of 3
    for (x in seq(3, timepoints, 3)) {
      start_time <- Sys.time()

      fit <- sem(generate_riclpm(x), data$data) # fit model

      converged    <- lavInspect(fit, "converged") # check convergence
      fit_measures <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli",
                                         "rmsea", "srmr"))
      coeff <- if (converged) coef(fit) else NULL

      # fill per condition results
      condition_results <- rbind(condition_results, data.frame(
        autoreg         = i,
        nconf           = j,
        timepoints_used = x,
        iteration       = x / 3,
        varRIx          = if (converged) coef(fit)["varRIx"] else NA,
        varRIy          = if (converged) coef(fit)["varRIy"] else NA,
        convergence     = converged,
        fit_measures    = I(list(fit_measures)),
        coefficients    = I(list(coeff)),
        processing_time = as.numeric(difftime(Sys.time(), start_time,
                                              units = "secs"))
      ))

      # aggressively remove fit to declutter memory
      rm(fit)
    }

    # add condition results to overall
    data_struct$results <- rbind(data_struct$results, condition_results)

    # save data to file
    filename <- save_clear(data_struct, filename_prefix)

    # message
    cat("Completed sim with autoregressive:", i, "and nconf:", j, "\n")

    rm(data)
    gc()
  }
}
```

#### Plotting

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggh4x) # Required for individual facet scaling

sim     <- Simulation1_20260311_212309 # load in data
results <- sim$results

results <- results %>%
  # unlist conditions
  mutate(
    autoreg = unlist(autoreg),
    nconf   = unlist(nconf)
  ) %>%
  # 
  mutate(
    bx = sapply(coefficients, function(c)
      if (!is.null(c) && "bx" %in% names(c)) c["bx"] else NA_real_)
  ) %>%
  filter(convergence == TRUE)

# true parameter value
true_beta <- 0.1

# One row per condition x timepoint: compute bias and keep varRIx
plot_data <- results %>%
  transmute(
    autoreg,
    nconf,
    timepoints_used,
    bias_bx = bx - true_beta,
    varRIx
  ) %>%
  pivot_longer(
    cols      = c(bias_bx, varRIx),
    names_to  = "outcome",
    values_to = "value"
  ) %>%
  mutate(
    outcome = factor(outcome,
                     levels = c("bias_bx", "varRIx"),
                     labels = c("Bias in bx", "Var(RI_x) estimate")),
    facet_row = factor(paste0("\u03b1 = ", autoreg)),
    facet_col = factor(paste0("Confounders = ", nconf),
                       levels = paste0("Confounders = ", sort(unique(nconf))))
  )

outcome_colours   <- c("Bias in bx"         = "#2166AC",
                        "Var(RI_x) estimate" = "#B2182B")
outcome_linetypes <- c("Bias in bx"         = "solid",
                        "Var(RI_x) estimate" = "dashed")

# ==============================================================================
# THEME
# ==============================================================================

theme_apa <- theme_classic(base_size = 11, base_family = "sans") +
  theme(
    axis.title         = element_text(size = 10),
    axis.text          = element_text(size = 9),
    axis.line          = element_line(colour = "black", linewidth = 0.4),
    strip.background   = element_blank(),
    strip.text         = element_text(face = "bold", size = 9),
    legend.position    = "bottom",
    legend.title       = element_blank(),
    legend.text        = element_text(size = 9),
    legend.key.width   = unit(1.5, "lines"),
    legend.key.size    = unit(0.5, "lines"),
    panel.grid.major.y = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin        = margin(8, 10, 6, 8),
    plot.caption       = element_text(size = 8, hjust = 0, margin = margin(t = 6))
  )

# ==============================================================================
# PLOT
# ==============================================================================

# 1. Update the data processing with the new labels
plot_data <- results %>%
  transmute(
    autoreg,
    nconf,
    timepoints_used,
    bias_bx = bx - true_beta,
    varRIx
  ) %>%
  pivot_longer(
    cols      = c(bias_bx, varRIx),
    names_to  = "outcome",
    values_to = "value"
  ) %>%
  mutate(
    outcome = factor(outcome,
                     levels = c("bias_bx", "varRIx"),
                     labels = c("Bias in cross-lagged parameter", 
                                "Random intercept variance")),
    facet_row = factor(paste0("\u03b1 = ", autoreg)),
    facet_col = factor(paste0("Confounders = ", nconf),
                       levels = paste0("Confounders = ", sort(unique(nconf))))
  )

outcome_colours <- c("Bias in cross-lagged parameter" = "#2166AC",
                     "Random intercept variance"      = "#B2182B")
outcome_linetypes <- c("Bias in cross-lagged parameter" = "solid",
                       "Random intercept variance"      = "dashed")

# 2. Define the emphasis data and scales
emphasis_box <- data.frame(facet_row = "α = 0.3", facet_col = "Confounders = 6")
shared_y_scale  <- scale_y_continuous(limits = c(0, 0.2)) 
special_y_scale <- scale_y_continuous(limits = c(0, 1.5))

# ==============================================================================
# PLOT
# ==============================================================================

p <- ggplot(plot_data,
            aes(x        = timepoints_used,
                y        = value,
                colour   = outcome,
                linetype = outcome)) +
  
  # The Callout Block (Background shading for bottom-right)
  geom_rect(data = emphasis_box, 
            aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
            fill = "grey80", alpha = 0.4, inherit.aes = FALSE) +

  geom_hline(yintercept = 0, linetype = "dotted",
             colour = "black", linewidth = 0.4) +
  
  geom_line(linewidth = 0.75, na.rm = TRUE) +
  
  facet_wrap(~ facet_row + facet_col, nrow = 2, scales = "free_y") + 
  
  facetted_pos_scales(
    y = list(
      facet_row == "α = 0.2" ~ shared_y_scale, 
      facet_col == "Confounders = 4" & facet_row == "α = 0.3" ~ shared_y_scale,
      facet_col == "Confounders = 6" & facet_row == "α = 0.3" ~ special_y_scale
    )
  ) +
  
  scale_colour_manual(values = outcome_colours) +
  scale_linetype_manual(values = outcome_linetypes) +
  scale_x_continuous(breaks = seq(0, 75, 15),
                     expand = expansion(mult = c(0.01, 0.02))) +
  
  theme_apa +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.key.width = unit(2.5, "lines") # Slightly wider for longer labels
  ) +
  
  labs(
    x       = "Number of timepoints",
    y       = "Value",
    caption = "Note. The shaded panel indicates a different Y-axis scale (0 to 1.5) compared to other panels (0 to 0.2).\n α = autoregressive effects of all variables, confounders = number of confounders on each variable of interest.\n "
  )

p

# ggsave("fig_sim1_2x2.png", p, width = 6.5, height = 5.5, dpi = 300)
```

### Simulation 2: the SRI-CLPM

Demonstrate usability of SRI-CLPM, so realistic sample sizes, set default DGM parameters. Just time-varying confounding, T = 3.

Plot idea, show 5 timepoints and plot the cross-lagged parameters with boxplots, with a line for the true value.

Total time it took is 276 mins (4.6 hours)

```{r}
library(lavaan)
library(parallel)

source("functionSource.R")

# ==============================================================================
# PARAMETERS
# ==============================================================================

timepoints   <- seq(6, 30, 6)
replications <- 3000
N_conditions <- c(250, 500)
timespans    <- c(2, 3)
seed         <- 42

true_bx <- 0.1
true_by <- 0.1

# ==============================================================================
# MODEL SPEC GENERATION
# CLPM and RI-CLPM: one spec per timepoint (no timespan dimension)
# SRI-CLPM variants: one spec per timepoint x timespan
# ==============================================================================

modelSpecs_clpm <- setNames(
  lapply(timepoints, function(i) generate_clpm(T = i)),
  as.character(timepoints)
)

modelSpecs_riclpm <- setNames(
  lapply(timepoints, function(i) generate_riclpm(T = i)),
  as.character(timepoints)
)

# Named list: modelSpecs_sriclpm[["ts2"]][["6"]], modelSpecs_sriclpm[["ts3"]][["12"]], etc.
modelSpecs_sriclpm <- setNames(
  lapply(timespans, function(ts)
    setNames(
      lapply(timepoints, function(i)
        generate_segmented_riclpm(T = i, timespan = ts, across_seg_cov = "toeplitz")),
      as.character(timepoints)
    )
  ),
  paste0("ts", timespans)
)

modelSpecs_sriclpm_free <- setNames(
  lapply(timespans, function(ts)
    setNames(
      lapply(timepoints, function(i)
        generate_segmented_riclpm(T = i, timespan = ts, across_seg_cov = "free")),
      as.character(timepoints)
    )
  ),
  paste0("ts", timespans)
)

# ==============================================================================
# HELPER: fit one model, return a single-row data frame
# ==============================================================================

fit_and_extract <- function(model_spec, model_label, tp, ts, N, data_df,
                            rep_id, true_bx, true_by) {

  fit <- tryCatch(sem(model_spec, data = data_df), error = function(e) NULL)

  if (!is.null(fit)) {
    converged  <- lavInspect(fit, "converged")
    post_check <- tryCatch(lavInspect(fit, "post.check"), error = function(e) FALSE)
    n_iter     <- tryCatch(fit@optim$iterations,           error = function(e) NA)

    coef_vec <- if (converged) coef(fit) else NULL
    se_vec   <- if (converged) tryCatch(sqrt(diag(vcov(fit))),
                                        error = function(e) NULL) else NULL
    fit_meas <- tryCatch(
      fitMeasures(fit, c("chisq", "df", "pvalue",
                         "cfi", "tli", "rmsea", "srmr", "aic", "bic")),
      error = function(e) rep(NA, 9)
    )
    rm(fit)
  } else {
    converged  <- FALSE
    post_check <- FALSE
    n_iter     <- NA
    coef_vec   <- NULL
    se_vec     <- NULL
    fit_meas   <- rep(NA, 9)
  }

  g <- function(vec, nm) {
    if (!is.null(vec) && nm %in% names(vec)) vec[nm] else NA_real_
  }

  data.frame(
    replication  = rep_id,
    N            = N,
    timepoints   = tp,
    timespan     = ts,       # NA for CLPM and RI-CLPM
    model_type   = model_label,

    convergence  = converged,
    post_check   = post_check,
    n_iterations = n_iter,

    bx      = g(coef_vec, "bx"),
    by      = g(coef_vec, "by"),
    bx_se   = g(se_vec,   "bx"),
    by_se   = g(se_vec,   "by"),
    bias_bx = g(coef_vec, "bx") - true_bx,
    bias_by = g(coef_vec, "by") - true_by,

    ax    = g(coef_vec, "ax"),
    ay    = g(coef_vec, "ay"),
    ax_se = g(se_vec,   "ax"),
    ay_se = g(se_vec,   "ay"),

    varRIx    = g(coef_vec, "varRIx"),
    varRIy    = g(coef_vec, "varRIy"),
    varRIx_se = g(se_vec,   "varRIx"),
    varRIy_se = g(se_vec,   "varRIy"),
    covRI     = g(coef_vec, "covRI"),
    covRI_se  = g(se_vec,   "covRI"),

    chisq  = fit_meas[1],
    df     = fit_meas[2],
    pvalue = fit_meas[3],
    cfi    = fit_meas[4],
    tli    = fit_meas[5],
    rmsea  = fit_meas[6],
    srmr   = fit_meas[7],
    aic    = fit_meas[8],
    bic    = fit_meas[9],

    stringsAsFactors = FALSE,
    row.names        = NULL
  )
}

# ==============================================================================
# WORKER: one full replication (all timepoints x timespans x models) for a given N
# ==============================================================================

run_replication <- function(rep_id, timepoints, timespans, seed, N,
                            true_bx, true_by,
                            modelSpecs_clpm, modelSpecs_riclpm,
                            modelSpecs_sriclpm, modelSpecs_sriclpm_free) {

  data <- DGM(N = N, timepoints = max(timepoints), seed = seed + rep_id)

  # Rows per rep: per timepoint — CLPM + RI-CLPM (once) + 2 timespans x 2 SRI variants
  n_rows_per_rep <- length(timepoints) * (2L + length(timespans) * 2L)
  rows <- vector("list", n_rows_per_rep)
  k    <- 1L

  for (i in timepoints) {
    tp_key <- as.character(i)
    data_i <- data$data[, c(paste0("x", 1:i), paste0("y", 1:i))]

    # CLPM and RI-CLPM — fitted once per timepoint, timespan = NA
    rows[[k]] <- fit_and_extract(modelSpecs_clpm[[tp_key]],
                                 "CLPM", i, NA_integer_, N,
                                 data_i, rep_id, true_bx, true_by)
    k <- k + 1L

    rows[[k]] <- fit_and_extract(modelSpecs_riclpm[[tp_key]],
                                 "RI-CLPM", i, NA_integer_, N,
                                 data_i, rep_id, true_bx, true_by)
    k <- k + 1L

    # SRI-CLPM variants — fitted once per timespan
    for (ts in timespans) {
      ts_key <- paste0("ts", ts)

      rows[[k]] <- fit_and_extract(modelSpecs_sriclpm[[ts_key]][[tp_key]],
                                   "SRI-CLPM", i, ts, N,
                                   data_i, rep_id, true_bx, true_by)
      k <- k + 1L

      rows[[k]] <- fit_and_extract(modelSpecs_sriclpm_free[[ts_key]][[tp_key]],
                                   "SRI-CLPM-free", i, ts, N,
                                   data_i, rep_id, true_bx, true_by)
      k <- k + 1L
    }

    rm(data_i)
  }

  rm(data)
  do.call(rbind, rows)
}

# ==============================================================================
# PARALLELISATION SETUP
# ==============================================================================

n_cores <- max(1L, detectCores() - 1L)
cat("Using", n_cores, "cores\n")

cl <- makeCluster(n_cores)

clusterExport(cl, varlist = c(
  "DGM", "generate_clpm", "generate_riclpm", "generate_segmented_riclpm",
  "fit_and_extract", "run_replication",
  "timepoints", "timespans", "seed", "true_bx", "true_by",
  "modelSpecs_clpm", "modelSpecs_riclpm",
  "modelSpecs_sriclpm", "modelSpecs_sriclpm_free"
))
clusterEvalQ(cl, { library(lavaan); library(mvtnorm); source("functionSource.R") })

# ==============================================================================
# MAIN LOOP — one parLapply per N condition
# ==============================================================================

data_struct      <- create_metadata()
start_time_total <- Sys.time()
all_results      <- vector("list", length(N_conditions))

for (ni in seq_along(N_conditions)) {
  N <- N_conditions[ni]
  cat("\n--- N =", N, "---\n")

  clusterExport(cl, varlist = "N", envir = environment())

  condition_results <- parLapply(
    cl,
    seq_len(replications),
    fun                     = run_replication,
    timepoints              = timepoints,
    timespans               = timespans,
    seed                    = seed,
    N                       = N,
    true_bx                 = true_bx,
    true_by                 = true_by,
    modelSpecs_clpm         = modelSpecs_clpm,
    modelSpecs_riclpm       = modelSpecs_riclpm,
    modelSpecs_sriclpm      = modelSpecs_sriclpm,
    modelSpecs_sriclpm_free = modelSpecs_sriclpm_free
  )

  all_results[[ni]] <- do.call(rbind, condition_results)

  elapsed <- round(difftime(Sys.time(), start_time_total, units = "mins"), 1)
  cat("N =", N, "done |", elapsed, "min elapsed\n")
  saveRDS(all_results[[ni]],
          paste0("sim_T_effect_checkpoint_N", N, ".rds"))
}

stopCluster(cl)

# ==============================================================================
# ASSEMBLE & SAVE
# ==============================================================================

replication_results <- do.call(rbind, all_results)
rownames(replication_results) <- NULL

data_struct$results <- replication_results
data_struct$simulation_parameters <- list(
  replications = replications,
  N_conditions = N_conditions,
  timepoints   = timepoints,
  timespans    = timespans,
  seed         = seed,
  true_bx      = true_bx,
  true_by      = true_by
)

filename <- save_clear(data_struct, filename_prefix = "sim_T_effect")

total_time <- difftime(Sys.time(), start_time_total, units = "mins")
cat("\n=== Simulation Completed ===\n")
cat("Total time:", round(total_time, 2), "minutes\n")
cat("Results saved to:", filename, "\n")
```

#### Plotting

```{r}
library(ggplot2)
library(dplyr)

sim     <- sim_T_effect_20260329_135008
results <- sim$results

true_beta <- 0.1

# Create a combined model label that folds timespan into the model name
# for SRI variants, and leaves CLPM/RI-CLPM as-is
plot_data <- results %>%
  mutate(
    model_label = case_when(
      model_type %in% c("CLPM", "RI-CLPM") ~ model_type,
      TRUE ~ paste0(model_type, " (SL=", timespan, ")")
    ),
    model_label = factor(model_label, levels = c(
      "CLPM",
      "RI-CLPM",
      "SRI-CLPM (SL=2)",
      "SRI-CLPM (SL=3)",
      "SRI-CLPM-free (SL=2)",
      "SRI-CLPM-free (SL=3)"
    )),
    N_label = factor(paste0("N = ", N),
                     levels = paste0("N = ", sort(unique(N))))
  )

mean_lines <- plot_data %>%
  group_by(model_label, timepoints, N_label) %>%
  summarise(mean_bx = mean(bx, na.rm = TRUE), .groups = "drop")

model_colours <- c(
  "CLPM"                 = "#636363",
  "RI-CLPM"              = "#2166AC",
  "SRI-CLPM (SL=2)"      = "#B2182B",
  "SRI-CLPM (SL=3)"      = "#F4A582",
  "SRI-CLPM-free (SL=2)" = "#4DAC26",
  "SRI-CLPM-free (SL=3)" = "#B8E186"
)

theme_apa <- theme_classic(base_size = 11, base_family = "sans") +
  theme(
    axis.title         = element_text(size = 10),
    axis.text          = element_text(size = 9),
    axis.line          = element_line(colour = "black", linewidth = 0.4),
    legend.position    = "bottom",
    legend.title       = element_blank(),
    legend.text        = element_text(size = 9),
    legend.key.size    = unit(0.5, "lines"),
    panel.grid.major.y = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin        = margin(8, 10, 6, 8),
    plot.caption       = element_text(size = 8, hjust = 0, margin = margin(t = 6))
  )

# 6 models — evenly spaced offsets centred on each timepoint tick
n_models     <- length(levels(plot_data$model_label))
dodge_offset <- 2.0
box_width    <- dodge_offset * 0.13

offsets <- setNames(
  seq(-dodge_offset / 2, dodge_offset / 2, length.out = n_models),
  levels(plot_data$model_label)
)

plot_data <- plot_data %>%
  mutate(x_dodged = timepoints + offsets[as.character(model_label)])

mean_lines <- mean_lines %>%
  mutate(x_dodged = timepoints + offsets[as.character(model_label)])

# Build one plot per N condition
make_plot <- function(n_val) {

  pd <- filter(plot_data, N_label == paste0("N = ", n_val))
  ml <- filter(mean_lines, N_label == paste0("N = ", n_val))

  ggplot(mapping = aes(colour = model_label, fill = model_label)) +

    geom_hline(yintercept = true_beta, linetype = "dashed",
               colour = "black", linewidth = 0.45) +

    geom_boxplot(
      data          = pd,
      aes(x = x_dodged, y = bx,
          group = interaction(timepoints, model_label)),
      width         = box_width,
      alpha         = 0.3,
      outlier.size  = 0.5,
      outlier.alpha = 0.2,
      linewidth     = 0.35
    ) +

#    geom_line(
#      data      = ml,
#      aes(x = x_dodged, y = mean_bx, group = model_label),
#      linewidth = 0.75
#    ) +

    scale_colour_manual(values = model_colours, drop = FALSE) +
    scale_fill_manual(values   = model_colours, drop = FALSE) +
    scale_x_continuous(
      breaks = unique(pd$timepoints),
      labels = unique(pd$timepoints)
    ) +
    labs(
      title   = paste0("N = ", n_val),
      x       = "Number of timepoints",
      y       = "Cross-lagged estimate",
      caption = if (n_val == max(sim$simulation_parameters$N_conditions)) {
        paste0(
          "Note. Boxes show interquartile range across replications; ",
          "Dashed reference line = true value. ",
          "SL = Segment Length. ",
          sim$simulation_parameters$replications, " replications per condition."
        )
      } else {
        NULL
      }
    ) +
    theme_apa +
    guides(colour = guide_legend(nrow = 2),
           fill   = guide_legend(nrow = 2))
}

p_list <- lapply(sort(unique(results$N)), make_plot)

# Print separately
for (p in p_list) print(p)

# ggsave("fig_T_effect_N250.png", p_list[[1]], width = 8, height = 5, dpi = 300)
# ggsave("fig_T_effect_N500.png", p_list[[2]], width = 8, height = 5, dpi = 300)
# ggsave("fig_T_effect_bx.png", p, width = 9, height = 4.5, dpi = 300)

# 1. Define True Values (Ensure true_ax and true_ay match your DGM settings)
true_ax <- 0.2  # Default autocorr_effects in your DGM is 0.2
true_ay <- 0.2

# 2. Calculate Coverage
coverage_summary <- results %>%
  filter(convergence == TRUE) %>%
  mutate(
    # Cross-lagged coverage
    bx_covered = (bx - 1.96 * bx_se) <= true_bx & (bx + 1.96 * bx_se) >= true_bx,
    by_covered = (by - 1.96 * by_se) <= true_by & (by + 1.96 * by_se) >= true_by,
    
    # Autoregressive coverage
    ax_covered = (ax - 1.96 * ax_se) <= true_ax & (ax + 1.96 * ax_se) >= true_ax,
    ay_covered = (ay - 1.96 * ay_se) <= true_ay & (ay + 1.96 * ay_se) >= true_ay
  ) %>%
  group_by(model_type, N, timepoints, timespan) %>%
  summarise(
    n_valid = n(),
    coverage_bx = mean(bx_covered, na.rm = TRUE),
    coverage_by = mean(by_covered, na.rm = TRUE),
    coverage_ax = mean(ax_covered, na.rm = TRUE),
    coverage_ay = mean(ay_covered, na.rm = TRUE),
    .groups = "drop"
  )

library(ggplot2)
library(tidyr)
library(dplyr)

# Filter for benchmarks and the specific SRI variant
heatmap_comparison_data <- coverage_summary %>%
  filter(
    model_type == "CLPM" | 
    model_type == "RI-CLPM" | 
    (model_type == "SRI-CLPM" & timespan == 2)
  ) %>%
  pivot_longer(
    cols = starts_with("coverage_"), 
    names_to = "Parameter", 
    values_to = "Rate"
  ) %>%
  mutate(
    # Explicit Parameter Labels
    Parameter_Label = case_when(
      Parameter == "coverage_ax" ~ "Autoregressive (X→X)",
      Parameter == "coverage_ay" ~ "Autoregressive (Y→Y)",
      Parameter == "coverage_bx" ~ "Cross-Lagged (X→Y)",
      Parameter == "coverage_by" ~ "Cross-Lagged (Y→X)"
    ),
    # Simplify model names for the plot headers
    Model_Label = factor(model_type, levels = c("CLPM", "RI-CLPM", "SRI-CLPM")),
    N_Label = paste0("N = ", N)
  )

ggplot(heatmap_comparison_data, aes(x = Parameter_Label, y = factor(timepoints), fill = Rate)) +
  geom_tile(color = "white") +
  # Add 2-decimal rounded values
  geom_text(aes(label = sprintf("%.2f", Rate)), color = "black", size = 2.8) +
  # Facet by Sample Size and Model
  facet_grid(N_Label ~ Model_Label) + 
  scale_fill_gradient(
    low = "red", 
    high = "deepskyblue3", 
    limits = c(0, 1),
    name = "Coverage"
  ) +
  theme_minimal() +
  labs(
    title = "95% CI Coverage Rates",
    subtitle = "Comparing CLPM & RI-CLPM vs. SRI-CLPM (Toeplitz, Segment length = 2)",
    x = "Model Parameter",
    y = "Number of Timepoints"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey95"),
    panel.spacing = unit(1, "lines")
  )
# 3. Display Results
view(coverage_summary)
```

### Explorative 2: Coverage rates

```{r}
library(future.apply)
source("functionSource.R")

N_conditions <- c(100, 150, 200)
replications <- 1000
timepoints   <- 20
segmentLength <- 2

# true values
true_bx <- 0.1
true_by <- 0.0

mdSpec <- generate_segmented_riclpm(timepoints, timespan = segmentLength)

# worker function
run_single_rep <- function(rep_idx, n_val) {

  data_list <- DGM(N = n_val, timepoints = timepoints,
                   beta_xy = true_bx, beta_yx = true_by,
                   seed = (42 + rep_idx + n_val))
  
  fit <- sem(mdSpec, data = data_list$data)
  
  pe <- parameterestimates(fit)
  
  # Extract unique values 
  res_bx <- pe %>% filter(label == "bx") %>% slice(1)
  res_by <- pe %>% filter(label == "by") %>% slice(1)
  
  return(data.frame(
    N      = n_val,
    rep    = rep_idx,
    bx_est = res_bx$est, bx_se = res_bx$se, bx_p = res_bx$pvalue,
    by_est = res_by$est, by_se = res_by$se, by_p = res_by$pvalue
  ))
}

plan(multisession, workers = parallel::detectCores() - 1)

cat("Starting parallel simulation...\n")

results_list <- future_lapply(N_conditions, function(n) {
  lapply(1:replications, function(r) run_single_rep(r, n)) %>% bind_rows()}, 
  future.seed = TRUE)

sim_results <- bind_rows(results_list)

save_clear(sim_results, filename_prefix = "Coverage_Simulation")

# analysis

sim_results <- Coverage_Simulation_20260501_003515

coverage_table <- sim_results %>%
  mutate(
    bx_covered = abs(bx_est - true_bx) / bx_se < 1.96,
    by_covered = abs(by_est - true_by) / by_se < 1.96
  ) %>%
  group_by(N) %>%
  summarise(
    Replications = n(),
    `Coverage bx (0.1)` = mean(bx_covered, na.rm = TRUE),
    `Coverage by (0.0)` = mean(by_covered, na.rm = TRUE),
    `Mean bx` = mean(bx_est),
    `Mean by` = mean(by_est)
  )

print(coverage_table)

# p-value plot

library(ggplot2)
library(dplyr)
library(tidyr)

# 1. Prepare data and calculate power/error rates for labels
power_labels <- sim_results %>%
  group_by(N) %>%
  summarise(
    bx_power = mean(bx_p < .05, na.rm = TRUE) * 100,
    by_error = mean(by_p < .05, na.rm = TRUE) * 100
  ) %>%
  pivot_longer(cols = c(bx_power, by_error), names_to = "parameter", values_to = "rate") %>%
  mutate(
    # Create the "N = X" label for faceting
    N_label = paste0("N = ", N),
    # Match the parameter labels used in the main plot data
    parameter = ifelse(parameter == "bx_power", "βx (True = 0.1)", "βy (True = 0.0)"),
    # Create the label text for the geom_text overlay
    label_text = paste0(ifelse(grepl("βx", parameter), "Power: ", "Type I Error: "), 
                        round(rate, 1), "%")
  )

# 2. Main Plot Data
plot_p_data <- sim_results %>%
  select(N, bx_p, by_p) %>%
  pivot_longer(cols = c(bx_p, by_p), names_to = "parameter", values_to = "p_val") %>%
  mutate(
    # Create the "N = X" label for faceting to match power_labels
    N_label = factor(paste0("N = ", N), levels = paste0("N = ", sort(unique(N)))),
    parameter = ifelse(parameter == "bx_p", "βx (True = 0.1)", "βy (True = 0.0)")
  )

# 3. Generate Plot
ggplot(plot_p_data, aes(x = p_val)) +
  geom_histogram(breaks = seq(0, 1, 0.05), fill = "#2166AC", colour = "white", alpha = 0.8) +
  
  # Add the Power/Error rate labels
  geom_text(data = power_labels, 
            aes(x = 0.5, y = Inf, label = label_text), 
            vjust = 2, size = 3.5, fontface = "bold", inherit.aes = FALSE) +
  
  # Use N_label instead of N for the grid rows
  facet_grid(N_label ~ parameter) +
  theme_classic() +
  labs(
    title = "Distribution of p-values by true parameter value and sample size",
    subtitle = "Power & Type I error: α = .05.",
    x = "p-value",
    y = "Frequency"
  )
```

### Simulation 3: A global intercept / time-invariant confounding

```{r}
library(lavaan)
library(parallel)

source("functionSource.R")

# ==============================================================================
# PARAMETERS
# ==============================================================================

DGMconditions <- c("only time-varying", "both")
replications  <- 1000
timepoints    <- 30
N             <- 200
seed          <- 42
timespan      <- 3

true_bx <- 0.1
true_by <- 0.1

# Generate model specs once — they are the same for every replication
modelSpec_sri    <- generate_segmented_riclpm(T = timepoints, timespan = timespan)
modelSpec_sri_g  <- generate_segmented_riclpm_global(T = timepoints, timespan = timespan)

# ==============================================================================
# HELPER: fit one model, return a single-row data frame
# ==============================================================================

fit_and_extract <- function(model_spec, model_label, dgm_label,
                            data_df, rep_id, true_bx, true_by) {

  fit <- tryCatch(
    sem(model_spec, data = data_df),
    error = function(e) NULL
  )

  if (!is.null(fit)) {
    converged  <- lavInspect(fit, "converged")
    post_check <- tryCatch(lavInspect(fit, "post.check"), error = function(e) FALSE)
    n_iter     <- tryCatch(fit@optim$iterations,           error = function(e) NA)

    coef_vec <- if (converged) coef(fit) else NULL
    se_vec   <- if (converged) tryCatch(sqrt(diag(vcov(fit))), error = function(e) NULL) else NULL

    fit_meas <- tryCatch(
      fitMeasures(fit, c("chisq", "df", "pvalue",
                         "cfi", "tli", "rmsea", "srmr", "aic", "bic")),
      error = function(e) rep(NA, 9)
    )
    rm(fit)
  } else {
    converged  <- FALSE
    post_check <- FALSE
    n_iter     <- NA
    coef_vec   <- NULL
    se_vec     <- NULL
    fit_meas   <- rep(NA, 9)
  }

  g <- function(vec, nm) {
    if (!is.null(vec) && nm %in% names(vec)) vec[nm] else NA_real_
  }

  data.frame(
    replication  = rep_id,
    dgm_condition = dgm_label,
    model_type   = model_label,

    convergence  = converged,
    post_check   = post_check,
    n_iterations = n_iter,

    # Cross-lagged parameters
    bx      = g(coef_vec, "bx"),
    by      = g(coef_vec, "by"),
    bx_se   = g(se_vec,   "bx"),
    by_se   = g(se_vec,   "by"),
    bias_bx = g(coef_vec, "bx") - true_bx,
    bias_by = g(coef_vec, "by") - true_by,

    # Autoregressive parameters
    ax    = g(coef_vec, "ax"),
    ay    = g(coef_vec, "ay"),
    ax_se = g(se_vec,   "ax"),
    ay_se = g(se_vec,   "ay"),

    # Segmented RI variances (both models)
    varRIx    = g(coef_vec, "varRIx"),
    varRIy    = g(coef_vec, "varRIy"),
    varRIx_se = g(se_vec,   "varRIx"),
    varRIy_se = g(se_vec,   "varRIy"),
    covRI     = g(coef_vec, "covRI"),
    covRI_se  = g(se_vec,   "covRI"),

    # Global RI variances (SRI-CLPM-G only; NA for standard SRI-CLPM)
    varRIgx    = g(coef_vec, "varRIgx"),
    varRIgy    = g(coef_vec, "varRIgy"),
    varRIgx_se = g(se_vec,   "varRIgx"),
    varRIgy_se = g(se_vec,   "varRIgy"),
    covRIg     = g(coef_vec, "covRIg"),
    covRIg_se  = g(se_vec,   "covRIg"),

    # Fit measures
    chisq  = fit_meas[1],
    df     = fit_meas[2],
    pvalue = fit_meas[3],
    cfi    = fit_meas[4],
    tli    = fit_meas[5],
    rmsea  = fit_meas[6],
    srmr   = fit_meas[7],
    aic    = fit_meas[8],
    bic    = fit_meas[9],

    stringsAsFactors = FALSE,
    row.names        = NULL
  )
}

# ==============================================================================
# WORKER: one replication for one DGM condition — called by parLapply
# ==============================================================================

run_replication <- function(rep_id, dgm_label, N, timepoints, seed,
                            true_bx, true_by,
                            modelSpec_sri, modelSpec_sri_g) {

  # Data generation
  dgm_args <- list(N = N, timepoints = timepoints, seed = seed + rep_id)
  if (dgm_label == "both") {
    dgm_args <- c(dgm_args, list(ri_var_x = 1, ri_var_y = 1, ri_cor = 0.1))
  }
  data <- do.call(DGM, dgm_args)

  row_sri   <- fit_and_extract(modelSpec_sri,   "SRI-CLPM",   dgm_label,
                               data$data, rep_id, true_bx, true_by)
  row_sri_g <- fit_and_extract(modelSpec_sri_g, "SRI-CLPM-G", dgm_label,
                               data$data, rep_id, true_bx, true_by)
  rm(data)

  rbind(row_sri, row_sri_g)
}

# ==============================================================================
# PARALLELISATION SETUP
# ==============================================================================

n_cores <- max(1L, detectCores() - 1L)
cat("Using", n_cores, "cores\n")

cl <- makeCluster(n_cores)

# Export everything workers need
clusterExport(cl, varlist = c(
  "DGM", "generate_segmented_riclpm", "generate_segmented_riclpm_global",
  "fit_and_extract", "run_replication",
  "modelSpec_sri", "modelSpec_sri_g",
  "N", "timepoints", "seed", "true_bx", "true_by"
))
clusterEvalQ(cl, { library(lavaan); library(mvtnorm); source("functionSource.R") })

# ==============================================================================
# MAIN LOOP — one parLapply per DGM condition
# ==============================================================================

data_struct       <- create_metadata()
start_time_total  <- Sys.time()
all_results       <- vector("list", length(DGMconditions))

for (ci in seq_along(DGMconditions)) {
  dgm_label <- DGMconditions[ci]
  cat("\n--- DGM condition:", dgm_label, "---\n")

  clusterExport(cl, varlist = "dgm_label", envir = environment())

  condition_results <- parLapply(
    cl,
    seq_len(replications),
    fun      = run_replication,
    dgm_label       = dgm_label,
    N               = N,
    timepoints      = timepoints,
    seed            = seed,
    true_bx         = true_bx,
    true_by         = true_by,
    modelSpec_sri   = modelSpec_sri,
    modelSpec_sri_g = modelSpec_sri_g
  )

  all_results[[ci]] <- do.call(rbind, condition_results)

  # Checkpoint after each condition completes
  elapsed <- round(difftime(Sys.time(), start_time_total, units = "mins"), 1)
  cat("Condition '", dgm_label, "' done |", elapsed, "min elapsed\n")
  saveRDS(all_results[[ci]],
          paste0("sim_srig_checkpoint_", gsub(" ", "_", dgm_label), ".rds"))
}

stopCluster(cl)

# ==============================================================================
# SAVE
# ==============================================================================

replication_results <- do.call(rbind, all_results)
rownames(replication_results) <- NULL

data_struct$results <- replication_results
data_struct$simulation_parameters <- list(
  replications  = replications,
  N             = N,
  timepoints    = timepoints,
  timespan      = timespan,
  DGMconditions = DGMconditions,
  seed          = seed,
  true_bx       = true_bx,
  true_by       = true_by
)

filename <- save_clear(data_struct, filename_prefix = "sim_srig_comparison")

total_time <- difftime(Sys.time(), start_time_total, units = "mins")
cat("\n=== Simulation Completed ===\n")
cat("Total time:", round(total_time, 2), "minutes\n")
cat("Results saved to:", filename, "\n")
```

#### Plotting

```{r}
sim     <- sim_srig_comparison_20260312_014452
results <- sim$results

true_beta <- 0.1

plot_data <- results %>%
  mutate(
    model_type    = factor(model_type, levels = c("SRI-CLPM", "SRI-CLPM-G")),
    dgm_condition = factor(dgm_condition,
                           levels = c("only time-varying", "both"),
                           labels = c("Only time-varying confounding",
                                      "Time-varying + time-invariant confounding"))
  )

model_colours <- c("SRI-CLPM"   = "#2166AC",
                   "SRI-CLPM-G" = "#B2182B")

theme_apa <- theme_classic(base_size = 11, base_family = "sans") +
  theme(
    axis.title         = element_text(size = 10),
    axis.text          = element_text(size = 9),
    axis.line          = element_line(colour = "black", linewidth = 0.4),
    strip.background   = element_blank(),
    strip.text         = element_text(face = "bold", size = 10),
    legend.position    = "bottom",
    legend.title       = element_blank(),
    legend.text        = element_text(size = 9),
    legend.key.size    = unit(0.5, "lines"),
    panel.grid.major.y = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin        = margin(8, 10, 6, 8),
    plot.caption       = element_text(size = 8, hjust = 0, margin = margin(t = 6))
  )

p <- ggplot(plot_data, aes(x = model_type, y = bx,
                            fill = model_type, colour = model_type)) +
  geom_hline(yintercept = true_beta, linetype = "dashed",
             colour = "black", linewidth = 0.45) +
  geom_boxplot(alpha        = 0.3,
               width        = 0.5,
               outlier.size  = 0.6,
               outlier.alpha = 0.3,
               linewidth     = 0.4) +
  facet_wrap(~dgm_condition) +
  scale_fill_manual(values   = model_colours) +
  scale_colour_manual(values = model_colours) +
  scale_x_discrete(labels = NULL) +
  labs(
    x       = NULL,
    y       = "Cross-lagged estimate (bx)",
    caption = paste0("Note. Dashed line = true value (", true_beta, "). ",
                     "N = ", sim$simulation_parameters$N, "; ",
                     sim$simulation_parameters$replications,
                     " replications per condition.")
  ) +
  theme_apa

p

# ggsave("fig_srig_comparison.png", p, width = 6.5, height = 4, dpi = 300)
```

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)

replication_results <- sim_srig_comparison_20260312_014452$results

# 1. Prepare the data: Pivot to wide format so we have one column per model
plot_data <- replication_results %>%
  # Filter for converged models only to ensure valid comparisons
  filter(convergence == TRUE) %>%
  # Keep only necessary columns
  select(replication, dgm_condition, model_type, bx) %>%
  # Pivot so model types are columns
  pivot_wider(
    names_from = model_type, 
    values_from = bx
  )

# 2. Create the scatterplot
ggplot(plot_data, aes(x = `SRI-CLPM`, y = `SRI-CLPM-G`)) +
  geom_point(size = 2, alpha = 0.1, color = "steelblue") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  # Separate plots for 'only time-varying' vs 'both'
  facet_wrap(~dgm_condition) +
  # Add the true parameter value as reference lines
  geom_vline(xintercept = 0.1, linetype = "dotted") +
  geom_hline(yintercept = 0.1, linetype = "dotted") +
  labs(
    title = "Cross-lagged Parameter Estimates Across Models",
    subtitle = "Dashed red line = Equal estimates; Dotted lines = True value (0.1)",
    x = "SRI-CLPM Estimate",
    y = "SGRI-CLPM Estimate"
  ) +
  theme_minimal()
```

### Explorative 3: Confounding Variance Decomposition

```{r}
library(future.apply)
source("functionSource.R")

# set-up
timepoints   <- 20
N            <- 200
replications <- 500 
plan(multisession, workers = parallel::detectCores() - 1)

# model specs
specs <- list(
  CLPM  = generate_clpm(timepoints),
  RICLPM = generate_riclpm(timepoints),
  SRI   = generate_segmented_riclpm(timepoints, timespan = 2),
  SGRI  = generate_segmented_riclpm_global(timepoints, timespan = 2)
)

cond_table <- data.frame(
  cond_name = c("pure", "time_var", "time_invar", "combined"),
  n_conf    = c(0, 5, 0, 5),
  ri_var    = c(0, 0, 1, 1),
  ri_cor    = c(0, 0,.1,.1)
)

# Worker function
compare_models_rep <- function(rep_idx, cond_row) {
  
  # Generate Data
  sim_data <- DGM(N = N, timepoints = timepoints, 
                  n_x_confounders = cond_row$n_conf, 
                  n_y_confounders = cond_row$n_conf,
                  ri_var_x = cond_row$ri_var,
                  ri_var_y = cond_row$ri_var,
                  ri_cor = cond_row$ri_cor,
                  seed = 123 + rep_idx)$data
  
  results <- lapply(names(specs), function(m_name) {
    fit <- sem(specs[[m_name]], data = sim_data)

    fm <- fitMeasures(fit, c("logl", "df", "aic", "bic", "cfi", "tli", "rmsea", "srmr"))
    pe <- parameterEstimates(fit)
    
    # Helper to safely extract parameters by label
    get_v <- function(lab) {
      v <- pe$est[pe$label == lab]
      if(length(v) > 0) v[1] else NA_real_
    }
    
    data.frame(
      model     = m_name,
      condition = cond_row$cond_name,
      rep       = rep_idx,
      bx      = get_v("bx"),
      varRIgx = get_v("varRIgx"),
      varRIgy = get_v("varRIgy"),
      varRIx  = get_v("varRIx"),
      varRIy  = get_v("varRIy"),
      logl = fm["logl"], df = fm["df"], AIC = fm["aic"], BIC = fm["bic"],
      CFI = fm["cfi"], TLI = fm["tli"], RMSEA = fm["rmsea"], SRMR = fm["srmr"]
    )
  })
  return(bind_rows(results))
}

# run
all_comparisons <- list()

for (i in 1:nrow(cond_table)) {
  cat("Simulating Condition:", cond_table$cond_name[i], "\n")
  all_comparisons[[i]] <- future_lapply(1:replications, function(r) {
    compare_models_rep(r, cond_table[i, ])
  }, future.seed = TRUE) %>% bind_rows()
}

final_comp_df <- bind_rows(all_comparisons)

# Save results
save_clear(final_comp_df, filename_prefix = "SGRI_CLPM_Simulation")

final_df <- SGRI_CLPM_Simulation_20260503_031342

# 1. Define the desired order of conditions
ordered_conditions <- c("pure", "time_var", "time_invar", "combined")

# 2. Calculate the mean estimates for the SGRI model
summary_table <- final_df %>%
  filter(model == "SGRI") %>%
  group_by(condition) %>%
  summarise(
    Mean_varRIgx = mean(varRIgx, na.rm = TRUE),
    Mean_varRIx  = mean(varRIx, na.rm = TRUE),
    Mean_bx      = mean(bx, na.rm = TRUE)
  ) %>%
  # 3. Apply the specific row order
  mutate(condition = factor(condition, levels = ordered_conditions)) %>%
  arrange(condition)

# Display the results
print(summary_table)

```

# REWRITE

PLOT 1

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggh4x) # Required for individual facet scaling

sim     <- Simulation1_20260311_212309 # load in data
results <- sim$results

results <- results %>%
  # unlist conditions
  mutate(
    autoreg = unlist(autoreg),
    nconf   = unlist(nconf)
  ) %>%
  # 
  mutate(
    bx = sapply(coefficients, function(c)
      if (!is.null(c) && "bx" %in% names(c)) c["bx"] else NA_real_)
  ) %>%
  filter(convergence == TRUE)

# true parameter value
true_beta <- 0.1

# One row per condition x timepoint: compute bias and keep varRIx
plot_data <- results %>%
  transmute(
    autoreg,
    nconf,
    timepoints_used,
    bias_bx = bx - true_beta,
    varRIx
  ) %>%
  pivot_longer(
    cols      = c(bias_bx, varRIx),
    names_to  = "outcome",
    values_to = "value"
  ) %>%
  mutate(
    outcome = factor(outcome,
                     levels = c("bias_bx", "varRIx"),
                     labels = c("Bias in bx", "Var(RI_x) estimate")),
    facet_row = factor(paste0("\u03b1 = ", autoreg)),
    facet_col = factor(paste0("Confounders = ", nconf),
                       levels = paste0("Confounders = ", sort(unique(nconf))))
  )

# ==============================================================================
# SHARED APA THEME  (defined once; reused by all subsequent plots)
# ==============================================================================

theme_apa <- theme_classic(base_size = 11, base_family = "sans") +
  theme(
    axis.title         = element_text(size = 10),
    axis.text          = element_text(size = 9),
    axis.line          = element_line(colour = "black", linewidth = 0.4),
    strip.background   = element_blank(),
    strip.text         = element_text(face = "bold", size = 9),
    legend.position    = "bottom",
    legend.title       = element_blank(),
    legend.text        = element_text(size = 9),
    legend.key.width   = unit(1.5, "lines"),
    legend.key.size    = unit(0.5, "lines"),
    panel.grid.major.y = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.major.x = element_blank(),
    plot.margin        = margin(8, 10, 6, 8),
    plot.caption       = element_text(size = 8, hjust = 0, margin = margin(t = 6))
  )

# Okabe-Ito colourblind-safe two-colour pair used for outcome lines
outcome_colours <- c(
  "Bias in cross-lagged parameter" = "#0072B2",  # blue
  "Random intercept variance"      = "#D55E00"   # vermilion
)
outcome_linetypes <- c(
  "Bias in cross-lagged parameter" = "solid",
  "Random intercept variance"      = "dashed"
)

# ==============================================================================
# PLOT DATA
# ==============================================================================

plot_data <- results %>%
  transmute(
    autoreg,
    nconf,
    timepoints_used,
    bias_bx = bx - true_beta,
    varRIx
  ) %>%
  pivot_longer(
    cols      = c(bias_bx, varRIx),
    names_to  = "outcome",
    values_to = "value"
  ) %>%
  mutate(
    outcome = factor(outcome,
                     levels = c("bias_bx", "varRIx"),
                     labels = c("Bias in cross-lagged parameter",
                                "Random intercept variance")),
    facet_row = factor(paste0("\u03b1 = ", autoreg)),
    facet_col = factor(paste0("Confounders = ", nconf),
                       levels = paste0("Confounders = ", sort(unique(nconf))))
  )

# 2. Define the emphasis data and scales
emphasis_box <- data.frame(facet_row = "α = 0.3", facet_col = "Confounders = 6")
shared_y_scale  <- scale_y_continuous(limits = c(0, 0.2)) 
special_y_scale <- scale_y_continuous(limits = c(0, 1.5))

# ==============================================================================
# PLOT
# ==============================================================================

p <- ggplot(plot_data,
            aes(x        = timepoints_used,
                y        = value,
                colour   = outcome,
                linetype = outcome)) +
  
  # The Callout Block (Background shading for bottom-right)
  geom_rect(data = emphasis_box, 
            aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
            fill = "grey80", alpha = 0.4, inherit.aes = FALSE) +

  geom_hline(yintercept = 0, linetype = "dotted",
             colour = "black", linewidth = 0.4) +
  
  geom_line(linewidth = 0.75, na.rm = TRUE) +
  
  facet_wrap(~ facet_row + facet_col, nrow = 2, scales = "free_y") + 
  
  facetted_pos_scales(
    y = list(
      facet_row == "α = 0.2" ~ shared_y_scale, 
      facet_col == "Confounders = 4" & facet_row == "α = 0.3" ~ shared_y_scale,
      facet_col == "Confounders = 6" & facet_row == "α = 0.3" ~ special_y_scale
    )
  ) +
  
  scale_colour_manual(values = outcome_colours) +
  scale_linetype_manual(values = outcome_linetypes) +
  scale_x_continuous(breaks = seq(0, 75, 15),
                     expand = expansion(mult = c(0.01, 0.02))) +
  
  theme_apa +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    legend.key.width = unit(2.5, "lines") # Slightly wider for longer labels
  ) +
  
  labs(
    x       = "Number of timepoints",
    y       = "Estimate"
#    caption = "Note. \u03b1 = autoregressive effect; confounders = number of time-arying confounders per variable.\nThe shaded panel uses a different y-axis scale (0\u20131.5); all other panels use 0\u20130.2."
  )

p

ggsave("images/fig_sim1_2x2.png", p, width = 6.5, height = 5.5, dpi = 300)
```

Plot 2

```{r}
sim     <- sim_T_effect_20260329_135008
results <- sim$results

true_beta <- 0.1

# Create a combined model label that folds timespan into the model name
# for SRI variants, and leaves CLPM/RI-CLPM as-is
plot_data <- results %>%
  mutate(
    model_label = case_when(
      model_type %in% c("CLPM", "RI-CLPM") ~ model_type,
      TRUE ~ paste0(model_type, " (SL=", timespan, ")")
    ),
    model_label = factor(model_label, levels = c(
      "CLPM",
      "RI-CLPM",
      "SRI-CLPM (SL=2)",
      "SRI-CLPM (SL=3)",
      "SRI-CLPM-free (SL=2)",
      "SRI-CLPM-free (SL=3)"
    )),
    N_label = factor(paste0("N = ", N),
                     levels = paste0("N = ", sort(unique(N))))
  )

mean_lines <- plot_data %>%
  group_by(model_label, timepoints, N_label) %>%
  summarise(mean_bx = mean(bx, na.rm = TRUE), .groups = "drop")

model_colours <- c(
  "CLPM"                 = "#999999",  # grey
  "RI-CLPM"              = "#0072B2",  # blue
  "SRI-CLPM (SL=2)"      = "#D55E00",  # vermilion
  "SRI-CLPM (SL=3)"      = "#E69F00",  # orange
  "SRI-CLPM-free (SL=2)" = "#009E73",  # green
  "SRI-CLPM-free (SL=3)" = "#56B4E9"   # sky blue
)

# theme_apa already defined in Simulation 1 plotting chunk

# 6 models — evenly spaced offsets centred on each timepoint tick
n_models     <- length(levels(plot_data$model_label))
dodge_offset <- 2.0
box_width    <- dodge_offset * 0.13

offsets <- setNames(
  seq(-dodge_offset / 2, dodge_offset / 2, length.out = n_models),
  levels(plot_data$model_label)
)

plot_data <- plot_data %>%
  mutate(x_dodged = timepoints + offsets[as.character(model_label)])

mean_lines <- mean_lines %>%
  mutate(x_dodged = timepoints + offsets[as.character(model_label)])

p <- ggplot(plot_data,
            mapping = aes(colour = model_label, fill = model_label)) +

  geom_hline(yintercept = true_beta, linetype = "dashed",
             colour = "black", linewidth = 0.45) +

  geom_boxplot(
    aes(x = x_dodged, y = bx,
        group = interaction(timepoints, model_label)),
    width         = box_width,
    alpha         = 0.4,
    outlier.size  = 0.5,
    outlier.alpha = 0.2,
    linewidth     = 0.35
  ) +

  facet_wrap(~ N_label, ncol = 1) +
  
  scale_colour_manual(values = model_colours, drop = FALSE) +
  scale_fill_manual(values   = model_colours, drop = FALSE) +
  scale_x_continuous(
    breaks = sort(unique(plot_data$timepoints)),
    labels = sort(unique(plot_data$timepoints))
  ) +

  labs(
    x       = "Number of timepoints",
    y       = "Cross-lagged estimate (\u03b2x)",
  ) +

  theme_apa +
  guides(colour = guide_legend(nrow = 2),
         fill   = guide_legend(nrow = 2))

p

ggsave("images/fig_T_effect_bx.png", p, width = 8, height = 9, dpi = 300)

# 1. Define True Values (Ensure true_ax and true_ay match your DGM settings)
true_ax <- 0.2  # Default autocorr_effects in your DGM is 0.2
true_ay <- 0.2

# 2. Calculate Coverage
coverage_summary <- results %>%
  filter(convergence == TRUE) %>%
  mutate(
    # Cross-lagged coverage
    bx_covered = (bx - 1.96 * bx_se) <= true_bx & (bx + 1.96 * bx_se) >= true_bx,
    by_covered = (by - 1.96 * by_se) <= true_by & (by + 1.96 * by_se) >= true_by,
    
    # Autoregressive coverage
    ax_covered = (ax - 1.96 * ax_se) <= true_ax & (ax + 1.96 * ax_se) >= true_ax,
    ay_covered = (ay - 1.96 * ay_se) <= true_ay & (ay + 1.96 * ay_se) >= true_ay
  ) %>%
  group_by(model_type, N, timepoints, timespan) %>%
  summarise(
    n_valid = n(),
    coverage_bx = mean(bx_covered, na.rm = TRUE),
    coverage_by = mean(by_covered, na.rm = TRUE),
    coverage_ax = mean(ax_covered, na.rm = TRUE),
    coverage_ay = mean(ay_covered, na.rm = TRUE),
    .groups = "drop"
  )

library(ggplot2)
library(tidyr)
library(dplyr)

# Filter for benchmarks and the specific SRI variant
heatmap_comparison_data <- coverage_summary %>%
  filter(
    model_type == "CLPM" | 
    model_type == "RI-CLPM" | 
    (model_type == "SRI-CLPM" & timespan == 2)
  ) %>%
  pivot_longer(
    cols = starts_with("coverage_"), 
    names_to = "Parameter", 
    values_to = "Rate"
  ) %>%
  mutate(
    # Explicit Parameter Labels
    Parameter_Label = case_when(
      Parameter == "coverage_ax" ~ "Autoregressive (X→X)",
      Parameter == "coverage_ay" ~ "Autoregressive (Y→Y)",
      Parameter == "coverage_bx" ~ "Cross-Lagged (X→Y)",
      Parameter == "coverage_by" ~ "Cross-Lagged (Y→X)"
    ),
    # Simplify model names for the plot headers
    Model_Label = factor(model_type, levels = c("CLPM", "RI-CLPM", "SRI-CLPM")),
    N_Label = paste0("N = ", N)
  )

p_heatmap <- ggplot(heatmap_comparison_data,
                    aes(x = Parameter_Label, y = factor(timepoints), fill = Rate)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", Rate)), colour = "black", size = 2.8) +
  facet_grid(N_Label ~ Model_Label) +
  scale_fill_gradient2(
    low      = "#D55E00",   # vermilion  (low coverage)
    mid      = "#F0E442",   # yellow     (nominal ~.95)
    high     = "#0072B2",   # blue       (high coverage)
    midpoint = 0.5,
    limits   = c(0, 1),
    name     = "Coverage rate",
  ) +
  scale_x_discrete(labels = function(x) gsub(" ", "\n", x)) +
  labs(
    x       = "Parameter",
    y       = "Number of timepoints"
  ) +
  theme_apa +
  theme(
    axis.text.x      = element_text(angle = 30, hjust = 1, size = 8),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 9),
    panel.spacing    = unit(0.8, "lines"),
    legend.key.width = unit(1.5, "lines"),
    legend.position = "none"
  )

p_heatmap

ggsave("images/fig_coverage_heatmap.png", p_heatmap, width = 7, height = 5.5, dpi = 300)
# 3. Display Results
```

Explo 2

```{r}
# analysis

sim_results <- Coverage_Simulation_20260501_003515

coverage_table <- sim_results %>%
  mutate(
    bx_covered = abs(bx_est - true_bx) / bx_se < 1.96,
    by_covered = abs(by_est - true_by) / by_se < 1.96
  ) %>%
  group_by(N) %>%
  summarise(
    Replications = n(),
    `Coverage bx (0.1)` = mean(bx_covered, na.rm = TRUE),
    `Coverage by (0.0)` = mean(by_covered, na.rm = TRUE),
    `Mean bx` = mean(bx_est),
    `Mean by` = mean(by_est)
  )

print(coverage_table)

# p-value plot

library(ggplot2)
library(dplyr)
library(tidyr)

# 1. Prepare data and calculate power/error rates for labels
power_labels <- sim_results %>%
  group_by(N) %>%
  summarise(
    bx_power = mean(bx_p < .05, na.rm = TRUE) * 100,
    by_error = mean(by_p < .05, na.rm = TRUE) * 100
  ) %>%
  pivot_longer(cols = c(bx_power, by_error), names_to = "parameter", values_to = "rate") %>%
  mutate(
    # Create the "N = X" label for faceting
    N_label = paste0("N = ", N),
    # Match the parameter labels used in the main plot data
    parameter = ifelse(parameter == "bx_power", "βx (True = 0.1)", "βy (True = 0.0)"),
    # Create the label text for the geom_text overlay
    label_text = paste0(ifelse(grepl("βx", parameter), "Power: ", "Type I Error: "), 
                        round(rate, 1), "%")
  )

# 2. Main Plot Data
plot_p_data <- sim_results %>%
  select(N, bx_p, by_p) %>%
  pivot_longer(cols = c(bx_p, by_p), names_to = "parameter", values_to = "p_val") %>%
  mutate(
    # Create the "N = X" label for faceting to match power_labels
    N_label = factor(paste0("N = ", N), levels = paste0("N = ", sort(unique(N)))),
    parameter = ifelse(parameter == "bx_p", "βx (True = 0.1)", "βy (True = 0.0)")
  )

p_pval <- ggplot(plot_p_data, aes(x = p_val)) +
  geom_histogram(breaks = seq(0, 1, 0.05),
                 fill = "#0072B2", colour = "white", alpha = 0.8) +
  geom_text(data       = power_labels,
            aes(x = 0.5, y = Inf, label = label_text),
            vjust       = 2, size = 3.2, fontface = "bold",
            inherit.aes = FALSE) +
  facet_grid(N_label ~ parameter) +
  scale_x_continuous(
  breaks = seq(0, 1, 0.25),
  labels = function(x) ifelse(x %in% c(0, 1), as.integer(x), x),
  expand = expansion(mult = c(0, 0.02))
) +
  labs(
    x       = "p-value",
    y       = "Frequency",
  ) +
  theme_apa +
  theme(strip.text = element_text(face = "bold", size = 9))

p_pval

ggsave("images/fig_power.png", p_pval, width = 7, height = 5.5, dpi = 300)
```

Plot 3

```{r}
sim     <- sim_srig_comparison_20260312_014452
results <- sim$results

true_beta <- 0.1

plot_data <- results %>%
  mutate(
    model_type    = factor(model_type, levels = c("SRI-CLPM", "SRI-CLPM-G")),
    dgm_condition = factor(dgm_condition,
                           levels = c("only time-varying", "both"),
                           labels = c("Only time-varying confounding",
                                      "Time-varying + time-invariant confounding"))
  )

plot_data <- plot_data %>%
  mutate(
    model_type = factor(model_type,
                        levels = c("SRI-CLPM", "SRI-CLPM-G"),
                        labels = c("SRI-CLPM",  "SGRI-CLPM"))
  )

model_colours <- c(
  "SRI-CLPM"  = "#0072B2",
  "SGRI-CLPM" = "#D55E00"
)

# theme_apa already defined in Simulation 1 plotting chunk

p <- ggplot(plot_data, aes(x = model_type, y = bx,
                            fill = model_type, colour = model_type)) +
  geom_hline(yintercept = true_beta, linetype = "dashed",
             colour = "black", linewidth = 0.45) +
  geom_boxplot(alpha        = 0.3,
               width        = 0.5,
               outlier.size  = 0.6,
               outlier.alpha = 0.3,
               linewidth     = 0.4) +
  facet_wrap(~dgm_condition) +
  scale_fill_manual(values   = model_colours) +
  scale_colour_manual(values = model_colours) +
  scale_x_discrete(labels = NULL) +
  labs(
    x       = NULL,
    y       = "Cross-lagged parameter (\u03b2x)"
  ) +
  theme_apa

p
ggsave("images/fig_srig_comparison.png", p, width = 6.5, height = 4, dpi = 300)
```

plot 4

```{r}
library(tidyr)
library(dplyr)

replication_results <- sim_srig_comparison_20260312_014452$results

# theme_apa already defined in Simulation 1 plotting chunk

# Pivot to wide format so we have one column per model
plot_data <- replication_results %>%
  select(replication, dgm_condition, model_type, bx) %>%
  pivot_wider(names_from = model_type, values_from = bx)

p_scatter <- ggplot(plot_data, aes(x = `SRI-CLPM`, y = `SRI-CLPM-G`)) +
  geom_point(size = 1.5, alpha = 0.15, colour = "#0072B2") +
  geom_abline(intercept = 0, slope = 1,
              linetype = "dashed", colour = "black", linewidth = 0.45) +
  geom_vline(xintercept = 0.1, linetype = "dotted",
             colour = "black", linewidth = 0.35) +
  geom_hline(yintercept = 0.1, linetype = "dotted",
             colour = "black", linewidth = 0.35) +
  facet_wrap(~dgm_condition) +
  scale_x_continuous(expand = expansion(mult = 0.02)) +
  scale_y_continuous(expand = expansion(mult = 0.02)) +
  labs(
    x       = "SRI-CLPM estimate (\u03b2x)",
    y       = "SGRI-CLPM estimate (\u03b2x)",
  ) +
  theme_apa +
  theme(strip.text = element_text(face = "bold", size = 9))

p_scatter
ggsave("images/fig_scatter.png", p_scatter, width = 6.5, height = 4, dpi = 300)
```

Plot 5

```{r}
final_df <- SGRI_CLPM_Simulation_20260503_031342

# 1. Define the desired order of conditions
ordered_conditions <- c("pure", "time_var", "time_invar", "combined")

# 2. Calculate the mean estimates for the SGRI model
summary_table <- final_df %>%
  filter(model == "SGRI") %>%
  group_by(condition) %>%
  summarise(
    Mean_varRIgx = mean(varRIgx, na.rm = TRUE),
    Mean_varRIx  = mean(varRIx, na.rm = TRUE),
    Mean_bx      = mean(bx, na.rm = TRUE)
  ) %>%
  # 3. Apply the specific row order
  mutate(condition = factor(condition, levels = ordered_conditions)) %>%
  arrange(condition)

# Display the results
print(summary_table)
```
