---
title: "Thesis Simulations"
author: "A.P. Zweistra"
format: html
editor: visual
---

This is intended to be the final, complete file that contains all necessary code to reproduce the results presented in the thesis (in combination with the "functionSource.R" file.

### Initialize

```{r}
library(mvtnorm)
library(tidyverse)
library(lavaan)

# Load function source file
source("functionSource.R")

## Function definitions from Bailey, 2024
row <- function( M ){
    if( ! ( is.matrix(M) && dim(M)[1] == dim(M)[2] && dim(M)[1]>0 ) ){
        rowM <- NULL
    } else {
        F <- dim(M)[1]
        rowM <- matrix( NA, nrow=F^2, ncol=1)
        for ( r in (1:F) ) {
            for( c in (1:F) ) {
                rowM[ c+F*(r-1), 1 ] <- M[r,c]
            }
        }
    }
    return( rowM )
}
irow <- function( rowM ){
    if( ! ( !is.null(rowM) && ( is.matrix(rowM) || (is.array(rowM) && length( which( dim(rowM) > 1 ) ) == 1) ) && dim(rowM)[1]>0 && dim(rowM)[2]==1 ) ){
        M <- NULL
    } else {
        F <- sqrt( length( rowM ) )
        M <- matrix( NA, nrow=F, ncol=F )
        for ( r in (1:F) ) {
            for ( c in (1:F) ) {
                M[r,c] <- rowM[ c+F*(r-1) ]
            }
        }
    }
    return( M )
}
```

### Simulation 1: the performance of the RI-CLPM over time.

2x2 matrix with DGM params nconf(4,6) and ax(.2,.3). Illustrate role of "temporal stability" in Illusory Traits. Illustrate direct relationship between illusory random intercept variance and less biased estimation of cross-lagged params. Population matrices.

```{r}
autoreg <- c(.2,.3)
nconf <- c(4,6)

# general settings
timepoints <- 30
seed <- 42
filename_prefix <- "Simulation1" # sets name for saved output
data_struct <- create_metadata() # returns empty results with metadata

for (i in autoreg){
  for (j in nconf){

    # empty object per condition results
    condition_results <- data.frame(
        autocorr = integer(),
        timepoints_used = integer(),
        iteration = integer(),
        varRIx = numeric(),
        varRIy = numeric(),
        convergence = logical(),
        fit_measures = I(list()),  
        coefficients = I(list()),  
        processing_time = numeric()
      )
    
    # generate data
    data <- DGM(autocorr_effects = i,
                n_x_confounders = j,
                n_y_confounders = j,
                timepoints = timepoints,
                seed = seed)
    
    # fit models on multiples of 3
    for (x in seq(3,timepoints,3)){
      start_time <- Sys.time()
      
      # take subset of data, fit model
      dataslice <- data$data[1:x,]
      modelSpec <- generate_riclpm(x)
      fit <- sem(modelSpec, dataslice)
      
      converged <- lavInspect(fit, "converged")
      fit_measures <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli",
                                       "rmsea", "srmr"))
      
      coeff <- if(converged){coef(fit)}else{NULL}
      
      condition_results <- rbind(condition_results, data.frame(
        autocorr = conditions_autocorr[a],  
        timepoints_used = i,
        iteration = i/3,
        varRIx = if(converged) coef(fit)["varRIx"] else NA,
        varRIy = if(converged) coef(fit)["varRIy"] else NA,
        convergence = converged,
        fit_measures = I(list(fit_measures)),
        coefficients = I(list(coeff)),
        processing_time = as.numeric(difftime(Sys.time(), start_time, 
                                              units = "secs"))
      ))
      rm(fit) #remove fit object so it doesnt clog up memory
    }
    data_struct$results <- rbind(data_struct$results, condition_results)
    
    filename <- save_clear(data_struct, filename_prefix)
    
    cat("Completed Sim with autoregressive: ", i, " and nconf: ", j)
    
    rm(data)
    gc()
    
  }
}
```

### Simulation 2: the SRI-CLPM

```{r}

```

### Simulation 3: 

```{r}

```
