---
title: "Thesis Simulations"
author: "A.P. Zweistra"
format: html
editor: visual
---

This is intended to be the final, complete file that contains all necessary code to reproduce the results presented in the thesis (in combination with the "functionSource.R" file.

### Initialize

```{r}
library(mvtnorm)
library(tidyverse)
library(lavaan)

# Load function source file
source("functionSource.R")

## Function definitions from Bailey, 2024
row <- function( M ){
    if( ! ( is.matrix(M) && dim(M)[1] == dim(M)[2] && dim(M)[1]>0 ) ){
        rowM <- NULL
    } else {
        F <- dim(M)[1]
        rowM <- matrix( NA, nrow=F^2, ncol=1)
        for ( r in (1:F) ) {
            for( c in (1:F) ) {
                rowM[ c+F*(r-1), 1 ] <- M[r,c]
            }
        }
    }
    return( rowM )
}
irow <- function( rowM ){
    if( ! ( !is.null(rowM) && ( is.matrix(rowM) || (is.array(rowM) && length( which( dim(rowM) > 1 ) ) == 1) ) && dim(rowM)[1]>0 && dim(rowM)[2]==1 ) ){
        M <- NULL
    } else {
        F <- sqrt( length( rowM ) )
        M <- matrix( NA, nrow=F, ncol=F )
        for ( r in (1:F) ) {
            for ( c in (1:F) ) {
                M[r,c] <- rowM[ c+F*(r-1) ]
            }
        }
    }
    return( M )
}
```

### Simulation 1: the performance of the RI-CLPM over time.

2x2 matrix with DGM params nconf(4,6) and ax(.2,.3). Illustrate role of "temporal stability" in Illusory Traits. Illustrate direct relationship between illusory random intercept variance and less biased estimation of cross-lagged params. Population matrices.

```{r}
autoreg <- c(.2,.3)
nconf <- c(4,6)

# general settings
timepoints <- 30
seed <- 42
filename_prefix <- "Simulation1" # sets name for saved output
data_struct <- create_metadata() # returns empty results with metadata

for (i in autoreg){
  for (j in nconf){

    # empty object per condition results
    condition_results <- data.frame(
        autoreg = integer(),
        timepoints_used = integer(),
        iteration = integer(),
        varRIx = numeric(),
        varRIy = numeric(),
        convergence = logical(),
        fit_measures = I(list()),  
        coefficients = I(list()),  
        processing_time = numeric()
      )
    
    # generate data
    data <- DGM(autocorr_effects = i,
                n_x_confounders = j,
                n_y_confounders = j,
                timepoints = timepoints,
                seed = seed)
    
    # fit models on multiples of 3
    for (x in seq(3,timepoints,3)){
      start_time <- Sys.time()
      
      # fit model
      fit <- sem(generate_riclpm(x), data$data)
      
      converged <- lavInspect(fit, "converged")
      fit_measures <- fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli",
                                       "rmsea", "srmr"))
      
      coeff <- if(converged){coef(fit)}else{NULL}
      
      condition_results <- rbind(condition_results, data.frame(
        autoreg = i,  
        timepoints_used = x,
        iteration = x/3,
        varRIx = if(converged) coef(fit)["varRIx"] else NA,
        varRIy = if(converged) coef(fit)["varRIy"] else NA,
        convergence = converged,
        fit_measures = I(list(fit_measures)),
        coefficients = I(list(coeff)),
        processing_time = as.numeric(difftime(Sys.time(), start_time, 
                                              units = "secs"))
      ))
      rm(fit) #remove fit object so it doesnt clog up memory
    }
    data_struct$results <- rbind(data_struct$results, condition_results)
    
    filename <- save_clear(data_struct, filename_prefix)
    
    cat("Completed Sim with autoregressive: ", i, " and nconf: ", j, "\n")
    
    rm(data)
    gc()
    
  }
}
```

### Simulation 2: the SRI-CLPM

Demonstrate usability of SRI-CLPM, so realistic sample sizes, set default DGM parameters. Just time-varying confounding, T = 3.

Plot idea, show 5 timepoints and plot the cross-lagged parameters with boxplots, with a line for the true value.

Rough estimated runtime 9 hours.

```{r}
timepoints <- seq(6, 30, 6)
replications <- 2000
N <- 200
seed <- 42

true_bx <- 0.1
true_by <- 0.1

data_struct <- create_metadata()

replication_results <- data.frame(
  replication  = integer(),
  timepoints   = integer(),
  model_type   = character(),

  # Convergence and diagnostics
  convergence  = logical(),
  post_check   = logical(),
  n_iterations = integer(),

  # Cross-lagged parameters
  bx    = numeric(),
  by    = numeric(),
  bx_se = numeric(),
  by_se = numeric(),
  bias_bx = numeric(),
  bias_by = numeric(),

  # Autoregressive parameters
  ax    = numeric(),
  ay    = numeric(),
  ax_se = numeric(),
  ay_se = numeric(),

  # Random intercept variances
  varRIx    = numeric(),
  varRIy    = numeric(),
  varRIx_se = numeric(),
  varRIy_se = numeric(),
  covRI     = numeric(),

  # Fit measures
  chisq  = numeric(),
  df     = numeric(),
  pvalue = numeric(),
  cfi    = numeric(),
  tli    = numeric(),
  rmsea  = numeric(),
  srmr   = numeric(),
  aic    = numeric(),
  bic    = numeric(),

  stringsAsFactors = FALSE
)

start_time_total <- Sys.time()

for (rep in 1:replications) {

  # Generate dataset of max size needed, then iteratively fit models on subsets.
  data <- DGM(N = N,
              timepoints = max(timepoints),
              seed = seed + rep)

  # loop through timepoint conditions to fit models
  for (i in timepoints) {

    # Subset columns to the first i timepoints for x and y
    data_i <- data$data[, c(paste0("x", 1:i), paste0("y", 1:i))]

    # --- FIT RI-CLPM ---
    start_time_riclpm <- Sys.time()

    fit_riclpm <- tryCatch({
      sem(generate_riclpm(T = i), data = data_i)
    }, error = function(e) NULL)

    if (!is.null(fit_riclpm)) {
      converged_riclpm  <- lavInspect(fit_riclpm, "converged")
      post_check_riclpm <- tryCatch(lavInspect(fit_riclpm, "post.check"),
                                    error = function(e) FALSE)
      n_iter_riclpm     <- tryCatch(fit_riclpm@optim$iterations,
                                    error = function(e) NA)

      coef_riclpm <- if (converged_riclpm) coef(fit_riclpm) else NULL
      se_riclpm   <- if (converged_riclpm) {
        tryCatch(sqrt(diag(vcov(fit_riclpm))), error = function(e) NULL)
      } else NULL

      fit_meas_riclpm <- tryCatch(
        fitMeasures(fit_riclpm, c("chisq", "df", "pvalue",
                                  "cfi", "tli", "rmsea", "srmr", "aic", "bic")),
        error = function(e) rep(NA, 9)
      )
    } else {
      converged_riclpm  <- FALSE
      post_check_riclpm <- FALSE
      n_iter_riclpm     <- NA
      coef_riclpm       <- NULL
      se_riclpm         <- NULL
      fit_meas_riclpm   <- rep(NA, 9)
    }

    bx_riclpm    <- if (!is.null(coef_riclpm) && "bx"     %in% names(coef_riclpm)) coef_riclpm["bx"]     else NA
    by_riclpm    <- if (!is.null(coef_riclpm) && "by"     %in% names(coef_riclpm)) coef_riclpm["by"]     else NA
    ax_riclpm    <- if (!is.null(coef_riclpm) && "ax"     %in% names(coef_riclpm)) coef_riclpm["ax"]     else NA
    ay_riclpm    <- if (!is.null(coef_riclpm) && "ay"     %in% names(coef_riclpm)) coef_riclpm["ay"]     else NA
    bx_se_riclpm <- if (!is.null(se_riclpm)   && "bx"     %in% names(se_riclpm))   se_riclpm["bx"]       else NA
    by_se_riclpm <- if (!is.null(se_riclpm)   && "by"     %in% names(se_riclpm))   se_riclpm["by"]       else NA
    ax_se_riclpm <- if (!is.null(se_riclpm)   && "ax"     %in% names(se_riclpm))   se_riclpm["ax"]       else NA
    ay_se_riclpm <- if (!is.null(se_riclpm)   && "ay"     %in% names(se_riclpm))   se_riclpm["ay"]       else NA
    varRIx_riclpm    <- if (!is.null(coef_riclpm) && "varRIx" %in% names(coef_riclpm)) coef_riclpm["varRIx"] else NA
    varRIy_riclpm    <- if (!is.null(coef_riclpm) && "varRIy" %in% names(coef_riclpm)) coef_riclpm["varRIy"] else NA
    varRIx_se_riclpm <- if (!is.null(se_riclpm)   && "varRIx" %in% names(se_riclpm))   se_riclpm["varRIx"]   else NA
    varRIy_se_riclpm <- if (!is.null(se_riclpm)   && "varRIy" %in% names(se_riclpm))   se_riclpm["varRIy"]   else NA
    covRI_riclpm     <- if (!is.null(coef_riclpm) && "covRI"  %in% names(coef_riclpm)) coef_riclpm["covRI"]  else NA

    replication_results <- rbind(replication_results, data.frame(
      replication  = rep,
      timepoints   = i,
      model_type   = "RI-CLPM",
      convergence  = converged_riclpm,
      post_check   = post_check_riclpm,
      n_iterations = n_iter_riclpm,
      bx      = bx_riclpm,
      by      = by_riclpm,
      bx_se   = bx_se_riclpm,
      by_se   = by_se_riclpm,
      bias_bx = bx_riclpm - true_bx,
      bias_by = by_riclpm - true_by,
      ax      = ax_riclpm,
      ay      = ay_riclpm,
      ax_se   = ax_se_riclpm,
      ay_se   = ay_se_riclpm,
      varRIx    = varRIx_riclpm,
      varRIy    = varRIy_riclpm,
      varRIx_se = varRIx_se_riclpm,
      varRIy_se = varRIy_se_riclpm,
      covRI     = covRI_riclpm,
      chisq  = fit_meas_riclpm[1],
      df     = fit_meas_riclpm[2],
      pvalue = fit_meas_riclpm[3],
      cfi    = fit_meas_riclpm[4],
      tli    = fit_meas_riclpm[5],
      rmsea  = fit_meas_riclpm[6],
      srmr   = fit_meas_riclpm[7],
      aic    = fit_meas_riclpm[8],
      bic    = fit_meas_riclpm[9],
      stringsAsFactors = FALSE
    ))

    if (!is.null(fit_riclpm)) rm(fit_riclpm)

    # --- FIT SRI-CLPM ---
    start_time_sriclpm <- Sys.time()

    fit_sriclpm <- tryCatch({
      sem(generate_segmented_riclpm(T = i, timespan = 3), data = data_i)
    }, error = function(e) NULL)

    if (!is.null(fit_sriclpm)) {
      converged_sriclpm  <- lavInspect(fit_sriclpm, "converged")
      post_check_sriclpm <- tryCatch(lavInspect(fit_sriclpm, "post.check"),
                                     error = function(e) FALSE)
      n_iter_sriclpm     <- tryCatch(fit_sriclpm@optim$iterations,
                                     error = function(e) NA)

      coef_sriclpm <- if (converged_sriclpm) coef(fit_sriclpm) else NULL
      se_sriclpm   <- if (converged_sriclpm) {
        tryCatch(sqrt(diag(vcov(fit_sriclpm))), error = function(e) NULL)
      } else NULL

      fit_meas_sriclpm <- tryCatch(
        fitMeasures(fit_sriclpm, c("chisq", "df", "pvalue",
                                   "cfi", "tli", "rmsea", "srmr", "aic", "bic")),
        error = function(e) rep(NA, 9)
      )
    } else {
      converged_sriclpm  <- FALSE
      post_check_sriclpm <- FALSE
      n_iter_sriclpm     <- NA
      coef_sriclpm       <- NULL
      se_sriclpm         <- NULL
      fit_meas_sriclpm   <- rep(NA, 9)
    }

    bx_sriclpm    <- if (!is.null(coef_sriclpm) && "bx"     %in% names(coef_sriclpm)) coef_sriclpm["bx"]     else NA
    by_sriclpm    <- if (!is.null(coef_sriclpm) && "by"     %in% names(coef_sriclpm)) coef_sriclpm["by"]     else NA
    ax_sriclpm    <- if (!is.null(coef_sriclpm) && "ax"     %in% names(coef_sriclpm)) coef_sriclpm["ax"]     else NA
    ay_sriclpm    <- if (!is.null(coef_sriclpm) && "ay"     %in% names(coef_sriclpm)) coef_sriclpm["ay"]     else NA
    bx_se_sriclpm <- if (!is.null(se_sriclpm)   && "bx"     %in% names(se_sriclpm))   se_sriclpm["bx"]       else NA
    by_se_sriclpm <- if (!is.null(se_sriclpm)   && "by"     %in% names(se_sriclpm))   se_sriclpm["by"]       else NA
    ax_se_sriclpm <- if (!is.null(se_sriclpm)   && "ax"     %in% names(se_sriclpm))   se_sriclpm["ax"]       else NA
    ay_se_sriclpm <- if (!is.null(se_sriclpm)   && "ay"     %in% names(se_sriclpm))   se_sriclpm["ay"]       else NA
    varRIx_sriclpm    <- if (!is.null(coef_sriclpm) && "varRIx" %in% names(coef_sriclpm)) coef_sriclpm["varRIx"] else NA
    varRIy_sriclpm    <- if (!is.null(coef_sriclpm) && "varRIy" %in% names(coef_sriclpm)) coef_sriclpm["varRIy"] else NA
    varRIx_se_sriclpm <- if (!is.null(se_sriclpm)   && "varRIx" %in% names(se_sriclpm))   se_sriclpm["varRIx"]   else NA
    varRIy_se_sriclpm <- if (!is.null(se_sriclpm)   && "varRIy" %in% names(se_sriclpm))   se_sriclpm["varRIy"]   else NA
    covRI_sriclpm     <- if (!is.null(coef_sriclpm) && "covRI"  %in% names(coef_sriclpm)) coef_sriclpm["covRI"]  else NA

    replication_results <- rbind(replication_results, data.frame(
      replication  = rep,
      timepoints   = i,
      model_type   = "SRI-CLPM",
      convergence  = converged_sriclpm,
      post_check   = post_check_sriclpm,
      n_iterations = n_iter_sriclpm,
      bx      = bx_sriclpm,
      by      = by_sriclpm,
      bx_se   = bx_se_sriclpm,
      by_se   = by_se_sriclpm,
      bias_bx = bx_sriclpm - true_bx,
      bias_by = by_sriclpm - true_by,
      ax      = ax_sriclpm,
      ay      = ay_sriclpm,
      ax_se   = ax_se_sriclpm,
      ay_se   = ay_se_sriclpm,
      varRIx    = varRIx_sriclpm,
      varRIy    = varRIy_sriclpm,
      varRIx_se = varRIx_se_sriclpm,
      varRIy_se = varRIy_se_sriclpm,
      covRI     = covRI_sriclpm,
      chisq  = fit_meas_sriclpm[1],
      df     = fit_meas_sriclpm[2],
      pvalue = fit_meas_sriclpm[3],
      cfi    = fit_meas_sriclpm[4],
      tli    = fit_meas_sriclpm[5],
      rmsea  = fit_meas_sriclpm[6],
      srmr   = fit_meas_sriclpm[7],
      aic    = fit_meas_sriclpm[8],
      bic    = fit_meas_sriclpm[9],
      stringsAsFactors = FALSE
    ))

    if (!is.null(fit_sriclpm)) rm(fit_sriclpm)

  }

  rm(data)
  if (rep %% 100 == 0) gc(verbose = FALSE)

  if (!(rep / replications * 100 %% 5)) {
    cat("Simulating... ", rep / replications * 100, " %\n")
  }

}

# Save results
data_struct$results <- replication_results
data_struct$simulation_parameters <- list(
  replications = replications,
  N            = N,
  timepoints   = timepoints,
  timespan     = 3,
  seed         = seed,
  true_bx      = true_bx,
  true_by      = true_by
)

filename <- save_clear(data_struct, filename_prefix = "sim_T_effect")

total_time <- difftime(Sys.time(), start_time_total, units = "mins")
cat("\n=== Simulation Completed ===\n")
cat("Total time:", round(total_time, 2), "minutes\n")
cat("Results saved to:", filename, "\n")
```

### Simulation 3: 

Wat als er ook time-invariant confounding is? Wat als we een globaal RI toevoegen?

set DGM params, combined confounding,

Een plot zou handig zijn om te laten zien dat het model hetzelfde functioneert. Het toegevoegd nut van de globale RI is dan moeilijk vast te stellen, daarvoor zou je moeten kunnen laten zien dat de scheiding tussen time-varying en time-invariant confounding goed werkt. Is dit buiten de scope van de thesis? Hoe zou ik dit willen laten zien als ik alle tijd van de wereld had?

Een optie zou zijn om beide specificaties van de SRI-CLPM te fitten op twee verschillende DGM's, met in beide gevallen time-varying confounding, maar in het ene geval ook time-invariant confounding.

In ieder geval is het nuttig om te kunnen laten zien dat de SRI-CLPM met globaal RI ongeveer dezelfde eigenschappen heeft als de SRI-CLPM zonder. Het is alleen niet nuttig om twee keer hetzelfde te plotten / simuleren in verschillende plotjes, dus dat moet vermeden worden.

Een betere optie is om de performance van de SRI-CLPM te plotten en te zoeken naar een manier om zowel de globale RI varianties en cross-lag bias in hetzelfde plaatje te krijgen.

```{r}

```
