------------------------------------------------------------------------

---
title: "Research Report Simulations"
author: "A.P. Zweistra"
format: html
editor: visual
---

This is the accompanying R code for the research report. For this file to function the functionSource.R file also needs to be present in the working directory. This file contains the implementations of the actual simulations themselves and the visualisations, drawing on functions sourced from the source file.

## Simulation

### Initialisation

Load in functions from the functionSource file and load necessary packages.

```{r}
## random seed and packages 
set.seed(427)

library(mvtnorm)
library(psych)
library(ggplot2)
library(reshape)
library(gridExtra)
library(lavaan)
library(reshape2)
library(tidyverse)

## Function definitions from Bailey
row <- function( M ){
    if( ! ( is.matrix(M) && dim(M)[1] == dim(M)[2] && dim(M)[1]>0 ) ){
        rowM <- NULL
    } else {
        F <- dim(M)[1]
        rowM <- matrix( NA, nrow=F^2, ncol=1)
        for ( r in (1:F) ) {
            for( c in (1:F) ) {
                rowM[ c+F*(r-1), 1 ] <- M[r,c]
            }
        }
    }
    return( rowM )
}
irow <- function( rowM ){
    if( ! ( !is.null(rowM) && ( is.matrix(rowM) || (is.array(rowM) && length( which( dim(rowM) > 1 ) ) == 1) ) && dim(rowM)[1]>0 && dim(rowM)[2]==1 ) ){
        M <- NULL
    } else {
        F <- sqrt( length( rowM ) )
        M <- matrix( NA, nrow=F, ncol=F )
        for ( r in (1:F) ) {
            for ( c in (1:F) ) {
                M[r,c] <- rowM[ c+F*(r-1) ]
            }
        }
    }
    return( M )
}
```

### Simulation 1

```{r}
source("functionSource.R")

# Simulation parameters
autocorr_level <- 0.2  
min_timepoints <- 3    # Minimum for RI-CLPM identification
max_timepoints <- 45
timepoint_increment <- 3  # fit models every three timepoints
filename_prefix <- "RR_clpm_vs_riclpm_comparison"

# Create metadata structure
data_struct <- create_metadata()

# True parameter values from DGM
true_bx <- 0.1
true_by <- 0.1

cat("Starting CLPM vs RI-CLPM comparison simulation\n")

# Generate data once with maximum timepoints
data <- DGM(N = 20000, 
            autocorr_effects = autocorr_level,
            timepoints = max_timepoints)

# Initialize storage
condition_results <- data.frame(
  timepoints_used = integer(),
  model_type = character(),  
  varRIx = numeric(),
  varRIy = numeric(),
  bx = numeric(),
  by = numeric(),
  ax = numeric(),
  ay = numeric(),
  bias_bx = numeric(),
  bias_by = numeric(),
  convergence = logical(),
  post_check = logical(),
  fit_measures = I(list()),
  coefficients = I(list()),
  processing_time = numeric(),
  stringsAsFactors = FALSE
)

# Generate sequence of timepoints
timepoints_seq <- seq(min_timepoints, max_timepoints, by = timepoint_increment)

# Loop over timepoints
for (i in timepoints_seq) {
  cat("Analyzing", i, "timepoints...\n")
  cat("Starting CLPM \n")
  
  # --- FIT CLPM ---
  start_time_clpm <- Sys.time()
  
  modelSpec_clpm <- generate_clpm(i)
  
  fit_clpm <- tryCatch({
    sem(model = modelSpec_clpm, data = data$data,
        estimator = "MLR",
        optim.method = "nlminb",
        control = list(eval.max = 2000, iter.max = 1000))
  }, error = function(e) {
    cat("  CLPM ERROR:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(fit_clpm)) {
    converged_clpm <- lavInspect(fit_clpm, "converged")
    post_check_clpm <- tryCatch({
      lavInspect(fit_clpm, "post.check")
    }, error = function(e) {
      FALSE
    })
    
    fit_measures_clpm <- tryCatch({
      fitMeasures(fit_clpm, c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr"))
    }, error = function(e) {
      rep(NA, 7)
    })
    
    coeff_clpm <- if(converged_clpm) {coef(fit_clpm)} else {NULL}
    
    bx_clpm <- if(converged_clpm && "bx" %in% names(coeff_clpm)) coeff_clpm["bx"] else NA
    by_clpm <- if(converged_clpm && "by" %in% names(coeff_clpm)) coeff_clpm["by"] else NA
    ax_clpm <- if(converged_clpm && "ax" %in% names(coeff_clpm)) coeff_clpm["ax"] else NA
    ay_clpm <- if(converged_clpm && "ay" %in% names(coeff_clpm)) coeff_clpm["ay"] else NA
    
  } else {
    converged_clpm <- FALSE
    post_check_clpm <- FALSE
    fit_measures_clpm <- rep(NA, 7)
    coeff_clpm <- NULL
    bx_clpm <- NA
    by_clpm <- NA
    ax_clpm <- NA
    ay_clpm <- NA
  }
  
  condition_results <- rbind(condition_results, data.frame(
    timepoints_used = i,
    model_type = "CLPM",
    varRIx = NA,  # CLPM doesn't have random intercepts
    varRIy = NA,
    bx = bx_clpm,
    by = by_clpm,
    ax = ax_clpm,
    ay = ay_clpm,
    bias_bx = bx_clpm - true_bx,
    bias_by = by_clpm - true_by,
    convergence = converged_clpm,
    post_check = post_check_clpm,
    fit_measures = I(list(fit_measures_clpm)),
    coefficients = I(list(coeff_clpm)),
    processing_time = as.numeric(difftime(Sys.time(), start_time_clpm, 
                                          units = "secs")),
    stringsAsFactors = FALSE
  ))
  
  if (!is.null(fit_clpm)) rm(fit_clpm)

  cat("Starting RI-CLPM \n")
  # --- FIT RI-CLPM ---
  start_time_riclpm <- Sys.time()
  
  modelSpec_riclpm <- generate_riclpm(i)
  
  fit_riclpm <- tryCatch({
    sem(model = modelSpec_riclpm, data = data$data,
        estimator = "MLR",
        optim.method = "nlminb",
        control = list(eval.max = 2000, iter.max = 1000))
  }, error = function(e) {
    cat("  RI-CLPM ERROR:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(fit_riclpm)) {
    converged_riclpm <- lavInspect(fit_riclpm, "converged")
    post_check_riclpm <- tryCatch({
      lavInspect(fit_riclpm, "post.check")
    }, error = function(e) {
      FALSE
    })
    
    fit_measures_riclpm <- tryCatch({
      fitMeasures(fit_riclpm, c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr"))
    }, error = function(e) {
      rep(NA, 7)
    })
    
    coeff_riclpm <- if(converged_riclpm) {coef(fit_riclpm)} else {NULL}
    
    bx_riclpm <- if(converged_riclpm && "bx" %in% names(coeff_riclpm)) coeff_riclpm["bx"] else NA
    by_riclpm <- if(converged_riclpm && "by" %in% names(coeff_riclpm)) coeff_riclpm["by"] else NA
    ax_riclpm <- if(converged_riclpm && "ax" %in% names(coeff_riclpm)) coeff_riclpm["ax"] else NA
    ay_riclpm <- if(converged_riclpm && "ay" %in% names(coeff_riclpm)) coeff_riclpm["ay"] else NA
    varRIx_riclpm <- if(converged_riclpm && "varRIx" %in% names(coeff_riclpm)) coeff_riclpm["varRIx"] else NA
    varRIy_riclpm <- if(converged_riclpm && "varRIy" %in% names(coeff_riclpm)) coeff_riclpm["varRIy"] else NA
    
  } else {
    converged_riclpm <- FALSE
    post_check_riclpm <- FALSE
    fit_measures_riclpm <- rep(NA, 7)
    coeff_riclpm <- NULL
    bx_riclpm <- NA
    by_riclpm <- NA
    ax_riclpm <- NA
    ay_riclpm <- NA
    varRIx_riclpm <- NA
    varRIy_riclpm <- NA
  }
  
  condition_results <- rbind(condition_results, data.frame(
    timepoints_used = i,
    model_type = "RI-CLPM",
    varRIx = varRIx_riclpm,
    varRIy = varRIy_riclpm,
    bx = bx_riclpm,
    by = by_riclpm,
    ax = ax_riclpm,
    ay = ay_riclpm,
    bias_bx = bx_riclpm - true_bx,
    bias_by = by_riclpm - true_by,
    convergence = converged_riclpm,
    post_check = post_check_riclpm,
    fit_measures = I(list(fit_measures_riclpm)),
    coefficients = I(list(coeff_riclpm)),
    processing_time = as.numeric(difftime(Sys.time(), start_time_riclpm, 
                                          units = "secs")),
    stringsAsFactors = FALSE
  ))
  
  if (!is.null(fit_riclpm)) rm(fit_riclpm)
  gc(verbose = FALSE)
}

# Add results to main data structure
data_struct$results <- condition_results

# Save results
filename <- save_clear(data_struct, filename_prefix)

cat("\n=== CLPM vs RI-CLPM simulation completed! ===\n")
cat("Results saved to:", filename, "\n")

```

### Simulation 2

```{r}
source("functionSource.R")

# Simulation parameters
autocorr_level <- 0.2  # Default autocorrelation
timespan_values <- c(2, 4, 6)
min_timepoints <- 3
max_timepoints <- 30
filename_prefix <- "RR_MRICLPM_timespan"

# Create metadata structure
data_struct <- create_metadata()

# True parameter values from DGM
true_bx <- 0.1
true_by <- 0.1

cat("Starting MRI-CLPM Free simulation with varying timespans\n")
cat("Timespans to test:", paste(timespan_values, collapse = ", "), "\n\n")

# Generate data once with 30 timepoints
data <- DGM(N = 20000, 
            autocorr_effects = autocorr_level,
            timepoints = 30)

# Initialize storage
condition_results <- data.frame(
  timespan = integer(),
  timepoints_used = integer(),
  n_segments = integer(),
  model_type = character(),  
  varRIx = numeric(),
  varRIy = numeric(),
  bx = numeric(),
  by = numeric(),
  ax = numeric(),
  ay = numeric(),
  bias_bx = numeric(),
  bias_by = numeric(),
  convergence = logical(),
  post_check = logical(),
  fit_measures = I(list()),
  coefficients = I(list()),
  processing_time = numeric(),
  stringsAsFactors = FALSE
)

# Loop over timespan values
for (ts in timespan_values) {
  cat("\n=== Timespan = ", ts, " ===\n")
  
  # Generate sequence of timepoints that are multiples of this timespan
  # between min_timepoints and max_timepoints
  timepoints_seq <- seq(ts, max_timepoints, by = ts)
  timepoints_seq <- timepoints_seq[timepoints_seq >= 3]  # Ensure at least 3 timepoints
  
  cat("Timepoints to analyze:", paste(timepoints_seq, collapse = ", "), "\n")
  cat("Number of models to fit:", length(timepoints_seq), "\n")
  
  # Loop over timepoints
  for (i in timepoints_seq) {
    
    # Calculate number of segments for reporting
    n_seg <- ceiling(i / ts)
    
    cat("  Analyzing", i, "timepoints (", n_seg, "segments)...\n")
    
    # --- FIT MRI-CLPM with Free Covariances ---
    start_time <- Sys.time()
    
    modelSpec <- generate_segmented_riclpm(i, timespan = ts, 
                                           across_seg_cov = "free")
    
    fit <- tryCatch({
      sem(model = modelSpec, data = data$data,
          estimator = "MLR",
          optim.method = "nlminb",
          control = list(eval.max = 2000, iter.max = 1000))
    }, error = function(e) {
      cat("    ERROR at timepoints =", i, ":", e$message, "\n")
      return(NULL)
    })
    
    if (!is.null(fit)) {
      converged <- lavInspect(fit, "converged")
      post_check <- tryCatch({
        lavInspect(fit, "post.check")
      }, error = function(e) {
        FALSE
      })
      
      fit_measures <- tryCatch({
        fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli", "rmsea", "srmr"))
      }, error = function(e) {
        rep(NA, 7)
      })
      
      coeff <- if(converged) {coef(fit)} else {NULL}
      
      bx_est <- if(converged && "bx" %in% names(coeff)) coeff["bx"] else NA
      by_est <- if(converged && "by" %in% names(coeff)) coeff["by"] else NA
      ax_est <- if(converged && "ax" %in% names(coeff)) coeff["ax"] else NA
      ay_est <- if(converged && "ay" %in% names(coeff)) coeff["ay"] else NA
      
      # Sum RI variances across segments
      varRIx_sum <- if(converged) {
        ri_vars <- grep("^varRIx[0-9]+$", names(coeff), value = TRUE)
        if(length(ri_vars) > 0) sum(coeff[ri_vars]) else NA
      } else NA
      
      varRIy_sum <- if(converged) {
        ri_vars <- grep("^varRIy[0-9]+$", names(coeff), value = TRUE)
        if(length(ri_vars) > 0) sum(coeff[ri_vars]) else NA
      } else NA
      
    } else {
      # Model fitting failed
      converged <- FALSE
      post_check <- FALSE
      fit_measures <- rep(NA, 7)
      coeff <- NULL
      bx_est <- NA
      by_est <- NA
      ax_est <- NA
      ay_est <- NA
      varRIx_sum <- NA
      varRIy_sum <- NA
    }
    
    # Store results
    condition_results <- rbind(condition_results, data.frame(
      timespan = ts,
      timepoints_used = i,
      n_segments = n_seg,
      model_type = "MRI-CLPM-Free",
      varRIx = varRIx_sum,
      varRIy = varRIy_sum,
      bx = bx_est,
      by = by_est,
      ax = ax_est,
      ay = ay_est,
      bias_bx = bx_est - true_bx,
      bias_by = by_est - true_by,
      convergence = converged,
      post_check = post_check,
      fit_measures = I(list(fit_measures)),
      coefficients = I(list(coeff)),  # This saves ALL coefficients including RI covariances
      processing_time = as.numeric(difftime(Sys.time(), start_time, 
                                            units = "secs")),
      stringsAsFactors = FALSE
    ))
    
    # Clean up
    if (!is.null(fit)) rm(fit)
    gc(verbose = FALSE)
  }
  
  cat("Completed timespan", ts, "\n")
}

# Add results to main data structure
data_struct$results <- condition_results

# Save results
filename <- save_clear(data_struct, filename_prefix)

cat("\n=== MRI-CLPM Free simulation completed! ===\n")
cat("Results saved to:", filename, "\n")

```

### Simulation 3

```{r}
source("functionSource.R")

# Generate data once with maximum timepoints
max_timepoints <- 45
data <- DGM(N = 20000, timepoints = max_timepoints, autocorr_effects = 0.2)

cat("Generated data with", max_timepoints, "timepoints\n")
cat("Data dimensions:", dim(data$data), "\n\n")

# Initialize storage
results <- data.frame(
  timepoints_used = integer(),
  model_type = character(),
  ax = numeric(),
  ay = numeric(),
  bx = numeric(),
  by = numeric(),
  stringsAsFactors = FALSE
)

# Fit models on increasing number of timepoints
timepoints_seq <- seq(3, max_timepoints, by = 2)

for (tp in timepoints_seq) {
  cat("Fitting models with", tp, "timepoints...\n")
  
  # Subset data to first tp timepoints
  cols_to_use <- c(paste0("x", 1:tp), paste0("y", 1:tp))
  data_subset <- data$data[, cols_to_use]
  
  # --- Fit RI-CLPM ---
  modelSpec_riclpm <- generate_riclpm(tp)
  fit_riclpm <- sem(modelSpec_riclpm, data = data_subset)
  
  if (lavInspect(fit_riclpm, "converged")) {
    coef_riclpm <- coef(fit_riclpm)
    results <- rbind(results, data.frame(
      timepoints_used = tp,
      model_type = "RI-CLPM",
      ax = coef_riclpm["ax"],
      ay = coef_riclpm["ay"],
      bx = coef_riclpm["bx"],
      by = coef_riclpm["by"],
      stringsAsFactors = FALSE
    ))
  }
  
  # --- Fit MRI-CLPM Toeplitz ---
  modelSpec_toeplitz <- generate_segmented_riclpm(tp, timespan = 2, across_seg_cov = "toeplitz")
  fit_toeplitz <- sem(modelSpec_toeplitz, data = data_subset)
  
  if (lavInspect(fit_toeplitz, "converged")) {
    coef_toeplitz <- coef(fit_toeplitz)
    results <- rbind(results, data.frame(
      timepoints_used = tp,
      model_type = "MRI-CLPM-Toeplitz",
      ax = coef_toeplitz["ax"],
      ay = coef_toeplitz["ay"],
      bx = coef_toeplitz["bx"],
      by = coef_toeplitz["by"],
      stringsAsFactors = FALSE
    ))
  }
  
  # --- Fit MRI-CLPM Free ---
  modelSpec_free <- generate_segmented_riclpm(tp, timespan = 2, across_seg_cov = "free")
  fit_free <- sem(modelSpec_free, data = data_subset)
  
  if (lavInspect(fit_free, "converged")) {
    coef_free <- coef(fit_free)
    results <- rbind(results, data.frame(
      timepoints_used = tp,
      model_type = "MRI-CLPM-Free",
      ax = coef_free["ax"],
      ay = coef_free["ay"],
      bx = coef_free["bx"],
      by = coef_free["by"],
      stringsAsFactors = FALSE
    ))
  }
}

# Print results
print(results)

# Test at 30 timepoints specifically
cat("\n=== Results at 30 timepoints ===\n")
results_30 <- results[results$timepoints_used == 30, ]
print(results_30)
```

## Visualisation

### Sim 1

```{r}
source("functionSource.R")

dataSim1 <- readRDS("RR_clpm_vs_riclpm_comparison_20251205_205030.rds")

names(dataSim1$results)

plot_crosslag_bias(
  dataSim1$results, x_var = "timepoints_used", group_var = "model_type", 
  true_bx = .1, true_by = .1, 
  x_label = "Timepoints considered", group_label = "Model", 
  title_suffix = "CLPM vs RI-CLPM",xlim = c(3,45),ylim = c(-.05,.05))
```

```{r}
library(ggplot2)
library(dplyr)
dataSim1 <- readRDS("RR_clpm_vs_riclpm_comparison_20251205_205030.rds")

condition_results <- dataSim1$results

# 1. Prepare data with Panel Markers
plot_data <- condition_results %>%
  select(timepoints_used, model_type, bias_bx, varRIx) %>%
  pivot_longer(cols = c(bias_bx, varRIx), 
               names_to = "metric", 
               values_to = "value") %>%
  filter(!(model_type == "CLPM" & metric == "varRIx")) %>%
  mutate(metric = factor(metric, 
                         levels = c("bias_bx", "varRIx"),
                         labels = c("(a) Bias in Cross-Lagged Parameter", 
                                    "(b) Random Intercept Variance")))

# 2. Simplified Publication-Ready Plot
ggplot(plot_data, aes(x = timepoints_used, y = value, group = model_type)) +
  geom_hline(data = filter(plot_data, grepl("(a)", metric)), 
             aes(yintercept = 0), linetype = "solid", color = "gray90") +
  
  geom_line(aes(linetype = model_type), size = 0.6) +
  geom_point(aes(shape = model_type), size = 2) +
  
  facet_wrap(~metric, ncol = 1, scales = "free_y") +
  
  scale_shape_manual(values = c("CLPM" = 1, "RI-CLPM" = 16)) +
  scale_linetype_manual(values = c("CLPM" = "dashed", "RI-CLPM" = "solid")) +
  labs(x = "Number of Timepoints", y = NULL, 
       shape = "Model Type:", linetype = "Model Type:") +
  
  # APA Theme with explicit font overrides
  theme_classic(base_size = 12, base_family = "serif") +
  theme(
    # Explicitly setting family here ensures the (a) and (b) markers match
    strip.background = element_blank(),
    strip.text = element_text(
      family = "serif", 
      hjust = 0, 
      face = "bold", 
      size = 12
    ), 
    legend.position = "bottom",
    legend.text = element_text(family = "serif"),
    legend.title = element_text(family = "serif", face = "bold"),
    axis.text = element_text(family = "serif", color = "black"),
    axis.title = element_text(family = "serif")
  )

ggsave("RICLPMfigure.png", height = 6, width = 10)
```

```{r}
dataSim1 <- readRDS("RR_clpm_vs_riclpm_comparison_20251205_205030.rds")
data <- dataSim1$results

# Filter to converged RI-CLPM models only (since CLPM doesn't have varRIx)
riclpm_data <- data[data$model_type == "RI-CLPM" & data$convergence == TRUE, ]

# Create plot with two y-axes
plot_combined <- ggplot(data = riclpm_data, aes(x = timepoints_used)) +
  # Primary axis: varRIx
  geom_line(aes(y = varRIx, color = "varRIx"), linewidth = 1) +
  geom_point(aes(y = varRIx, color = "varRIx"), size = 2) +
  # Secondary axis: bias_bx (scaled to fit same range)
  geom_line(aes(y = bias_bx * max(varRIx, na.rm = TRUE) / max(abs(bias_bx), na.rm = TRUE), 
                color = "bias_bx"), linewidth = 1) +
  geom_point(aes(y = bias_bx * max(varRIx, na.rm = TRUE) / max(abs(bias_bx), na.rm = TRUE), 
                 color = "bias_bx"), size = 2) +
  # Add zero line for bias (scaled)
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", alpha = 0.5) +
  # Scales
  scale_y_continuous(
    name = "Random Intercept Variance (varRIx)",
    sec.axis = sec_axis(~ . * max(abs(riclpm_data$bias_bx), na.rm = TRUE) / 
                          max(riclpm_data$varRIx, na.rm = TRUE),
                        name = "Bias in bx (Y→X)")
  ) +
  scale_color_manual(values = c("varRIx" = "#0072B2", "bias_bx" = "#D55E00"),
                     labels = c("bias_bx" = "Bias in bx (Y→X)", 
                                "varRIx" = "Random Intercept Variance (X)")) +
  labs(title = "RI-CLPM: Random Intercept Variance and Cross-lagged Bias",
       x = "Number of Timepoints",
       color = "Parameter") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.position = "bottom",
        axis.title.y.left = element_text(color = "#0072B2"),
        axis.title.y.right = element_text(color = "#D55E00"))

print(plot_combined)
```

### Sim 2

```{r}
source("functionSource.R")

dataSim2 <- readRDS("RR_MRICLPM_timespan_20251205_212008.rds ")

names(dataSim2$results)

plot_crosslag_bias(
  dataSim2$results, x_var = "timepoints_used", group_var = "timespan", 
  true_bx = .1, true_by = .1, 
  x_label = "Timepoints considered", group_label = "Timespan", 
  title_suffix = "MRI-CLPM",xlim = c(3,30),ylim = c(-.025,.025))
```

```{r}
library(ggplot2)
library(dplyr)
source("functionSource.R")

dataSim2 <- readRDS("RR_MRICLPM_timespan_20251205_212008.rds ")

# 1. Prepare data
# Ensure timespan is a factor so ggplot treats it as discrete categories
plot_data_ts <- dataSim2$results %>%
  mutate(timespan_label = factor(timespan, 
                                 levels = c(2, 4, 6),
                                 labels = c("Timespan = 2", "Timespan = 4", "Timespan = 6")))

# 2. Simplified Publication-Ready Plot
ggplot(plot_data_ts, aes(x = timepoints_used, y = bias_bx, 
                         group = timespan_label, 
                         linetype = timespan_label, 
                         shape = timespan_label)) +
  # Reference line at 0 (No Bias)
  geom_hline(yintercept = 0, linetype = "solid", color = "gray90") +
  
  # Lines and Points
  geom_line(size = 0.6) +
  geom_point(size = 2.5, fill = "white") +
  ylim(c(-0.01,0.02))+
  
  # Labels
  labs(title = "Bias in Cross-Lagged Parameter by Segment Timespan",
       x = "Number of Timepoints",
       y = "Bias",
       linetype = "Segment Timespan:",
       shape = "Segment Timespan:") +
  
  # Grayscale scaling
  scale_linetype_manual(values = c("solid", "dashed", "dotdash")) +
  scale_shape_manual(values = c(16, 17, 15)) + # Circle, Triangle, Square
  
  # APA Theme with explicit font overrides
  theme_classic(base_size = 12, base_family = "serif") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13, family = "serif"),
    axis.title = element_text(face = "bold", family = "serif"),
    axis.text = element_text(color = "black", family = "serif"),
    legend.position = "bottom",
    legend.text = element_text(family = "serif"),
    legend.title = element_text(family = "serif", face = "bold"),
    axis.line = element_line(size = 0.5)
  ) 

ggsave("timespanFigure.png", width = 10, height = 6)
```

### Sim 3

```{r}
library(ggplot2)
library(dplyr)

# 1. Prepare the data: Filter for bx and calculate bias
plot_data <- results %>%
  mutate(bias_bx = bx - 0.1)

# 2. Generate the APA-style Grayscale Plot
plot <- ggplot(plot_data, aes(x = timepoints_used, y = bias_bx, 
                      group = model_type, 
                      shape = model_type, 
                      linetype = model_type)) +
  # Add a reference line at 0 (No Bias)
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +
  # Add lines and points in grayscale
  geom_line() +
  geom_point(size = 3, fill = "white") + # White fill helps shapes stand out on lines
  # Formatting scales for grayscale/shapes
  scale_shape_manual(values = c(16, 17, 15)) + # Circle, Triangle, Square
  scale_linetype_manual(values = c("solid", "dashed", "dotdash")) +
  # Labeling
  labs(title = "Bias in Cross-Lagged Parameter (X -> Y) Across Timepoints",
       x = "Number of Timepoints",
       y = "Bias (Estimate - 0.10)",
       shape = "Model Type",
       linetype = "Model Type") +
  # APA Theming
  theme_classic() + 
  theme(
    text = element_text(family = "serif", size = 12),      # Times New Roman equivalent
    plot.title = element_text(hjust = 0.5, face = "bold", size = 13),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",                            # Legend at bottom for journal space
    legend.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    axis.line = element_line(size = 0.5)
  ) +
  # Use exact decimals on Y axis for precision
  scale_y_continuous(labels = function(x) sprintf("%.3f", x))

ggsave("toeplitzFigure.png",width = 10, height = 6)
```

```{r}
# Load required libraries
if (!require("flextable")) install.packages("flextable")
library(flextable)
library(officer)

# 1. Create the data frame
table_data <- data.frame(
  Simulation = c("1", "2.1", "2.2"), # 4 rows as requested
  Aim = c("Extend findings by Bailey et al. (2024).", "Assess choice of segment timespan for the SRI-CLPM.", "Test Toeplitz constraint on latent variable covariance matrix in the SRI-CLPM"),
  stringsAsFactors = FALSE
)

# 2. Build the APA table
apa_table <- flextable(table_data) %>%
  # Set font to Serif (Times New Roman equivalent)
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 12, part = "all") %>%
  
  # Remove all default borders
  border_remove() %>%
  
  # Add APA standard horizontal lines (Top, Bottom, and Header)
  hline_top(border = fp_border(width = 1.5), part = "header") %>%
  hline_bottom(border = fp_border(width = 1.5), part = "header") %>%
  hline_bottom(border = fp_border(width = 1.5), part = "body") %>%
  
  # Formatting
  align(align = "left", part = "all") %>%
  autofit() %>%
  set_caption(caption = "Table 1. Overview of Simulation Conditions")

# 3. View the table
print(apa_table)

# Optional: Export to Word (standard for academic submissions)
 read_docx() %>%
   body_add_flextable(apa_table) %>%
   print(target = "simulation_table.docx")
```
